### 2012-08-28 b0-llvm.tex
Populating repository

\documentclass{article}

\usepackage[english]{babel}
\usepackage{amssymb,amsmath}
\usepackage[colorlinks=true]{hyperref}
\usepackage{framed}
\usepackage{a4wide}

\title{Translation of B implementations to the LLVM (\textsc{draft})}
\author{} \date{}

\newcommand{\trad}[2]{\ensuremath{\lVert \textsf{#1} \rVert^{\textit{#2}}}}
\newcommand{\nl}[0]{\ensuremath{\hookleftarrow}}
\newcommand{\mty}[0]{\texttt{""}}
\DeclareMathOperator{\conc}{\smallfrown}
\DeclareMathOperator{\isdef}{\blacktriangleright}
\DeclareMathOperator{\dom}{\mbox{dom}}
\DeclareMathOperator{\fresh}{\textit{fresh}}
\DeclareMathOperator{\name}{\mathcal{L}()}

\begin{document}

\maketitle

\begin{abstract}
  This document compiles ideas to produce software components to the
  LLVM intermediate format from B implementations. It is a work in
  progress and may be later used as a basis to produce a translator
  from B0 to LLVM.
\end{abstract}

[[[ TODO: PROOF READ ]]]

[[[ SYNTHESIZE TYPES OF PARSE TREE NODES AND THEIR ATTRIBUTES ]]]

\tableofcontents

\section{Introduction}
\label{sec:intro}

The goal of this work is to connect the B method to the LLVM
compilation framework. It is motivated by the fact that many compiling
technologies are implemented in the LLVM framework: optimization,
static analysis, code generation. Providing a connection from the B
method to this framework gives the users of the B method access to
these technologies. Our approach is to define a translation from B0
(the subset of the B language that is used to describe imperative
programs) to the LLVM intermediate format.

\section{Notations and basic definitions}
\label{sec:notations}

The translation is defined by structural recursion on the source
language B0. For each grammatical construct, a code generation
function is defined: $\trad{~}{op}$ for operations, $\trad{~}{inst}$
for instructions, $\trad{~}{exp}$ for expressions and so forth. Each
such function usually produces a text corresponding to bits of LLVM
source code (we use $text$ to denote the type for textual values). It
may have additional parameters and results. In particular, some of
this extra data is a total mapping from B0 identifiers and types to
LLVM text ($dict$ denotes the type of such mapping). Additional
information on those functions will be given below.

We assume that the B0 program has been parsed, and an attributed parse
tree of this program is available. Each node \textsf{n} in the tree
has an attribute \textsf{k} that classifies the node in a syntactic
category. For instance, if \textsf{n.k = BEq}, then \textsf{n}
represents a ``becomes assigned'' instruction. Moreover, for each kind
of node that may appear as an element of a list, there exists an
attribute \textsf{next} that points to the next node in that
list. This is the case for, e.g. instructions (in a block), branches
(in a conditional instruction), names (in a declaration list), etc.
For the last element of that list, the value of this attribute is
$\bot$. Table~\ref{tab:node-attr} summarizes the different kinds of
nodes and their attributes.

\begin{table}
  \begin{center}
    \begin{tabular}[t]{cc}
      \begin{tabular}[t]{|l|l|l|}
        \hline
        \textsf{k} & Decription & attributes\\
        \hline
        \hline
        \textsf{Impl} & Implementation & \textsf{sees} : ? \\
        & & \textsf{imports} : ? \\
        & & $\cdots$ \\
        & & \textsf{ops} : Op \\
        \hline
        \textsf{Op} & Operation & \textsf{header}: \textsf{Hd} \\
        & & $\textsf{body} \in \textsf{inst}$ \\
        & & \textsf{next} : \textsf{Op} \\
        \hline
        \textsf{Hd} & Operation header & \textsf{ident} : text \\
        & & \textsf{output} : \textsf{VarDecl} \\
        & & \textsf{input} : \textsf{VarDecl} \\
        \hline
        \textsf{Var} & Variable, parameter & \textsf{ident} : text \\
        & & \textsf{type} : basic type \\
        & & \textsf{next} : \textsf{Var} \\
        \hline
      \end{tabular}
      &
      \begin{tabular}[t]{|l|l|l|}
        \hline
        \textsf{k} & Decription & attributes\\
        \hline
        \hline
        \textsf{Blk} & Block instr. & $\textsf{body} \in \textsf{inst}$ \\
        & & $\textsf{next} \in \textsf{Instruction}$ \\
        \hline
        \textsf{VarDecl} & Variable decl. & \textsf{ident}: \textsf{VarDecl} \\
        & & $\textsf{body} \in \textsf{inst}$ \\
        \hline
        \textsf{If} & If & \textsf{branches}: \textsf{IfBr} \\
        & & $\textsf{next} \in \textsf{Instruction}$ \\
        \hline
        \textsf{IfBr} & If branch & \textsf{cond}: predicate \\
        & & $\textsf{body} \in \textsf{Instr}$ \\
        & & \textsf{next} : \textsf{IfBr} \\
        \hline
        \textsf{BEq} & Becomes equal & \textsf{lhs}: target \\
        & & \textsf{rhs}: expression \\
        & & $\textsf{next} \in \textsf{Instruction}$ \\
        \hline
        \textsf{Call} & Call instruction & \textsf{fun}: name \\
        & & \textsf{args}: expression \\
        & & $\textsf{next} \in \textsf{Instruction}$ \\
        \hline
        \textsf{While} & While instr. & \textsf{cond}: predicate \\
        & & $\textsf{body} \in \textsf{Instr}$ \\
        & & $\textsf{next} \in \textsf{Instruction}$ \\
        \hline
        \textsf{Assert} & Assert instr. & $\textsf{next} \in \textsf{Instruction}$ \\
        \hline
        \textsf{Case} & Case instr. & \textsf{expr}: expression\\
        & & \textsf{branches}: \textsf{CaseBr} \\
        & & $\textsf{next} \in \textsf{Instruction}$ \\
        \hline
        \textsf{CaseBr} & Case branch & \textsf{val}: \textsf{Value} \\
        & & $\textsf{body} \in \textsf{Instr}$ \\
        & & \textsf{next} : \textsf{CaseBr} \\
        \hline
        \textsf{Value} & Value & ??? \\
        & & \textsf{next}: \textsf{Value} \\
        \hline
      \end{tabular} \\
      \\
      \multicolumn{2}{c}{Note: $\textsf{Instr} = \{\textsf{Blk}, \textsf{VarDecl}, \textsf{If}, \textsf{BEq}, \textsf{Call}, \textsf{While}, \textsf{Assert}, \textsf{Case}\}$} \\
    \end{tabular}
    \caption{Structure of the parse tree nodes. For each attribute we give the name of the attribute and
    the type or kind(s) of node that it may be.}
    \label{tab:node-attr}
  \end{center}
\end{table}

Code generation is defined through a set of numbered rules. The
translation rules for some element \textsf{E} of the B0 language are
presented using a pseudo-functional language.  In such rules, the
\textsf{sans serif} and \texttt{typewriter} fonts are used to denote
\textsf{B0} abstract syntax tree nodes and \texttt{LLVM} text
respectively, and $\conc$ denotes string catenation and $\nl$ a new
line. .

Several utility functions are employed throughout those rules.  The
operator $\conc^,$ is the concatenation of two texts separated by a
comma---provided both texts are not empty.
\begin{align}
A \conc^, B \isdef & \textbf{ if } A = \texttt{""} \textbf{ then } B \nonumber \\
& \textbf{ elsif } B = \texttt{""} \textbf{ then } A \nonumber \\
& \textbf{ else } A \conc \texttt{", "} \conc B \nonumber
\end{align}

$\fresh: text \times dict \rightarrow text$ is such that
$\fresh(t, D)$  returns a text $t'$ that is either $t$, if $t$ is not in
the domain of the dictionary $D$, or $t$ followed by the string $n$ of
digits representing the smallest positive number $t \conc n$ does not
belong to the domain of $D$.

We assume that there is an unlimited supply of labels and local variables.
$\name$ returns a text that represents a fresh label or local variable
name for the LLVM currently being produced:
\begin{framed}
\begin{align}
\begin{split}
  &\name : \rightarrow text \\
  &\textbf{let } v : \mathbb{N} = 0 \textbf{ in let } \name = v := v + 1, \textbf{yields } (\textit{text } v)
\end{split}
\end{align}
\end{framed}

\section{A bird's-eye view of the translation}
\label{sec:view}

LLVM programs are structured in \emph{modules}. A LLVM module contains
global \emph{variables}, \emph{functions} and \emph{declarations} of
external entities. The role of variables is to carry some state,
functions may act on that state, and declarations specify the type of
of the external entities.

A B implementation shall then be mapped to a LLVM module.  We also
assume the existence of B library machines that model LLVM variables
that are of a basic type, equipped with operations to read and write
their value.  Other machines, corresponding to more complex types may
also be imported. One may map each section of the B implementation to
the LLVM as summarized in the following table:

\begin{center}
  \begin{tabular}{|c|c|}
    \hline
    B0 & LLVM \\
    \hline
    \hline
    implementation & module \\
    \hline
    imported library machines & global variables \\
    \hline
    other imported machines & other modules \\
    \hline
    operations & functions \\
    \hline
    external references & declarations \\
    \hline
  \end{tabular}
\end{center}

The definition of this translation is structured in four parts:
\begin{description}
\item[control] This part is responsible for describing how B
  operations, and more specifically, B0 instructions, can be mapped to
  LLVM statements.
\item[data] This part is responsible for describing how the data types
  compatible with B0 are translated to LLVM data types.
\item[expressions] This part describes how the evaluation of B0
  expressions and conditions may be codified in LLVM.
\item[modularity] This part is responsible for describing a strategy
  to translate separately the different B modules that may compose a
  B0 implementation to different LLVM source files.
\end{description}

We first define the translation rule for the root of the source
program abstract syntax tree, which shall correspond to a B
implementation. This is a preliminary version where the code
generation is only able to handle standalone B0 implementations, i.e.
the state is defined in the ``concrete variables'' clause. Moreover
the data is restricted to predefined basic types \textsf{INT} and
\textsf{BOOL}:
\begin{framed}
\begin{align}
\begin{split}
  \trad{}{impl} : & \textsf{ Node} \rightarrow text \\
  \trad{i}{impl} \isdef & \textbf{ let } O_{c} = \trad{n.values}{val} \textbf{ and } O_v = \trad{n.vars}{var} \\
  & \quad \textbf{ and } O_{in} = \trad{n.initialization}{ini} \textbf{ and } O_{op} = \trad{n.ops}{ops} \\
  & \quad \textbf{ in} \\
  & \textbf{ yields } \\
  & \quad \trad{i.ops}{ops}  \\
  & \textbf{ where } \textsf{i.k} = \textsf{Impl} 
\end{split}
\end{align}
\end{framed}
Code generation for initialisation and operations is specified in
section~\ref{sec:control}. Code generation for data (type and
state) is defined in section~\ref{sec:data}, and for expression
evaluation in section~\ref{sec:expr}.

\section{Data}
\label{sec:data}

\subsection{Type}
\label{sec:type}
For the time being, the translation only considers two types of
data: integers and booleans. In the case of integers, the translation
is based on the supposition that the B0 definition of \textsf{INT}
is the range $-2^{31}.. 2^{31}-1$:

\begin{framed}
\begin{align}
\begin{split}
  \trad{BOOL}{type} \isdef & \texttt{ "i1"} \\
  \trad{INT}{type} \isdef & \texttt{ "i32"}
\end{split}
\end{align}
\end{framed}

Future version of this document should handle other types of data,
namely integer intervals, enumerated sets, arrays and records (and
any other implementable data type that B0 allows).

\subsection{State}
\label{sec:state}

Concrete variables are mapped to so-called internal variables in the
LLVM program, i.e. private data that is local to the produced
file. The following rule describes the code generation for such
variables:
\begin{framed}
\begin{align}
\begin{split}
  \trad{}{var} : & \textsf{ Node} \times dict \rightarrow text \times dict \\
  \trad{v, $D$}{type} \isdef & \textbf{ let } O', D' = \trad{v.next, $D$}{var} \textbf{ and } 
  id = \texttt{"@"} \conc \textsf{v.ident} \textbf{ in} \\
  & \textbf{ yields } \\
  & \quad (id \conc \texttt{" = internal global "} \conc \trad{v.type}{type} \conc \nl \\
  & \quad \quad \conc O'), D' \cup \{ \textsf{n} \mapsto id \} \\
  & \textbf{ where } \textsf{v} \ne \bot \textbf{ and } \textsf{v.k} = \textsf{ConcVar} \\
  \trad{$\bot$, $D$}{type} \isdef & \textbf{ yields } \texttt{""}, D
\end{split}
\end{align}
\end{framed}

\subsection{Constant values}
\label{sec:values}

Constant valuations are translated to constants in the
LLVM program, as described in the following rule:
\begin{framed}
\begin{align}
\begin{split}
  \trad{}{val} : & \textsf{ Node} \times dict \rightarrow text \times dict \\
  \trad{v, $D$}{val} \isdef & \textbf{ let } O', D' = \trad{v.next, $D$}{var} \textbf{ and } 
  id = \texttt{"@"} \conc \textsf{v.ident} \textbf{ in} \\
  & \textbf{ yields } \\
  & \quad (id \conc \texttt{" = constant "} \conc \trad{v.type}{type} \conc \nl \\
  & \quad \quad \conc O'), D' \cup \{ \textsf{n} \mapsto id \} \\
  & \textbf{ where } \textsf{v} \ne \bot \textbf{ and } \textsf{v.k} = \textsf{ConcVar} \\
  \trad{$\bot$, $D$}{val} \isdef & \textbf{ yields } \texttt{""}, D
\end{split}
\end{align}
\end{framed}

\section{Expressions}
\label{sec:expr}

This section presents rules for the generation of LLVM code to
evaluate B0 expressions. These rules are restricted to expressions of
the following basic types: \textsf{BOOL} and \textsf{INT}. Extension
to handle other basic types, arrays and record is left for future
work.

The definition of the code generation is divided as follows:
section~\ref{sec:const} present the base cases (boolean and integer
literals, identifiers), section~\ref{sec:arith} presents code
generation rules for arithmetic expressions, section~\ref{sec:rel} and
section~\ref{sec:cond} is dedicated to relations and conditions.

\subsection{Constants and identifiers}
\label{sec:const}

The following definition provides a mapping from B0 boolean literals
to the corresponding LLVM values:
$$D_{bool} = \{ \textsf{FALSE} \mapsto \texttt{"0"}, \textsf{TRUE} \mapsto \texttt{"1"} \}.$$

Based on this definition, the translation of boolean literals is straightforward:
\begin{framed}
\begin{align}
\begin{split}
  \trad{n, $D$}{exp} \isdef & \textbf{ yields } \texttt{""}, D_{bool}(\textsf{n.val}), \texttt{"i1"} \\
  & \textbf{ where } \textsf{n.k} = \textsf{LitBool}
\end{split}
\end{align}
\end{framed}

The translation of integer literals is even simpler, as they have the
same denotation in B0 and LLVM:
\begin{framed}
\begin{align}
\begin{split}
  \trad{n, $D$}{exp} \isdef & \textbf{ yields } \texttt{""}, \textsf{n.val}, \texttt{"i32"} \\
  & \textbf{ where } \textsf{n.k} = \textsf{LitInt}
\end{split}
\end{align}
\end{framed}

The last such rule specifies the translation of identifiers. It mainly
consists in a lookup in the translation dictionary, and shall return
also the corresponding LLVM data type:

\begin{framed}
\begin{align}
\begin{split}
  \trad{n, $D$}{exp} \isdef & \textbf{ yields } \texttt{""}, D(\textsf{n}), \trad{n.type}{type} \\
  & \textbf{ where } \textsf{n.k} = \textsf{Ident}
\end{split}
\end{align}
\end{framed}

\subsection{Arithmetic operations}
\label{sec:arith}

The following definition maps B0 binary operators to LLVM binary statements:

$$D_{bari} = \{ + \mapsto \texttt{"add"}, 
- \mapsto \texttt{"sub"}, * \mapsto \texttt{"mul"}, / \mapsto
\texttt{"sdiv"}, \% \mapsto \texttt{"srem"} \}.$$

Based upon this definition, the following rule specifies the
generation of LLVM code responsible for the evaluation of binary
arithmetic expressions (plus the name of the local variable storing
the resulting value, and of its LLVM type):
\begin{framed}
\begin{align}
\begin{split}
  \trad{n, $D$}{exp} \isdef & \textbf{ let } O_p, O_l, O_t =
  \trad{n.left, $D$}{exp} \textbf{ and }
  O'_p, O'_l, O'_t = \trad{n.right, $D$}{exp} \\
  & \quad \textbf{ and } v = \mathcal{V} \textbf{ and } op = D_{bari}(\textsf{n.op}) \textbf{ in } \\
  & \textbf{ yields} \\
  & \quad (O_p \conc \\
  & \quad O'_p \conc \\
  & \quad v \conc \texttt{" = i32 "} \conc O_l \conc^, O'_l \conc \nl), \\
  & \quad v, \texttt{i32} \\
  & \textbf{ where } \textsf{n.k} = \textsf{Arith} \textbf{ and } \textsf{n.op} \in \dom D_{bari} \\
  & \textbf{ assert } O_t = O'_t = \texttt{i32}
\end{split}
\end{align}
\end{framed}

Next rule specifies the code generation for the successor and
predecessor operators. As LLVM has no corresponding statement, the
generated code uses addition and subtraction:
\begin{framed}
\begin{align}
\begin{split}
  \trad{n, $D$}{exp} \isdef & \textbf{ let } O_p, O_l, O_t =
  \trad{n.arg, $D$}{exp} \textbf{ and } v = \mathcal{V} \textbf{ in } \\
  & \textbf{ yields} \\
  & \quad (O_p \conc \\
  & \quad v \conc \texttt{" = i32 add 1, "} \conc O_l \conc \nl), \\
  & \quad v, \texttt{i32} \\
  & \textbf{ where } \textsf{n.k} = \textsf{Arith} \textbf{ and } \textsf{n.op} = \textsf{succ} \\
  & \textbf{ assert } O_t = \texttt{i32} \\
  \trad{n, $D$}{exp} \isdef & \textbf{ let } O_p, O_l, O_t =
  \trad{n.arg, $D$}{exp} \textbf{ and } v = \mathcal{V} \textbf{ in } \\
  & \textbf{ yields} \\
  & \quad (O_p \conc \\
  & \quad v \conc \texttt{" = i32 sub "} \conc O_l \conc \texttt{", 1"} \conc \nl), \\
  & \quad v, \texttt{i32} \\
  & \textbf{ where } \textsf{n.k} = \textsf{Arith} \textbf{ and } \textsf{n.op} = \textsf{pred} \\
  & \textbf{ assert } O_t = \texttt{i32}
\end{split}
\end{align}
\end{framed}

The following rule specifies code generation for the unary
minus. Since LLVM does not have a specific statement for such
operation, the translation uses subtraction:
\begin{framed}
\begin{align}
\begin{split}
  \trad{n, $D$}{exp} \isdef & \textbf{ let } O_p, O_l, O_t =
  \trad{n.arg, $D$}{exp} \textbf{ and } v = \mathcal{V} \textbf{ in } \\
  & \textbf{ yields} \\
  & \quad (O_p \conc \\
  & \quad v \conc \texttt{" = i32 sub 0, "} \conc O_l \conc \nl), \\
  & \quad v, \texttt{i32} \\
  & \textbf{ where } \textsf{n.k} = \textsf{Arith} \textbf{ and } \textsf{n.op} = \textsf{-} \\
  & \textbf{ assert } O_t = \texttt{i32}
\end{split}
\end{align}
\end{framed}

Also, LLVM does not have a predefined operation for the ``power''
operator provided by B0. The corresponding rule introduces a call
to a function, named \texttt{@power} that implements this operator (the LLVM source code
for such function is not presented herein):
\begin{framed}
\begin{align}
\begin{split}
  \trad{n, $D$}{rel} \isdef & \textbf{ let } O_p, O_l, O_t =
  \trad{n.left, $D$}{exp} \textbf{ and }
  O'_p, O'_l, O'_t = \trad{n.right, $D$}{exp} \\
  & \quad \textbf{ and } l = \mathcal{V} \textbf{ in } \\
  & \textbf{ yields} \\
  & \quad (O_p \conc \\
  & \quad O'_p \conc \\
  & \quad l \conc \texttt{" = call i32 @power(i32, "} \conc O_l \conc \texttt{", i32 "} \conc O'_l \conc \texttt{")"}) \conc \nl), \\
  & \quad l, \texttt{"i32"} \\
  & \textbf{ where } \textsf{r.k} = \textsf{Power} \\
  & \textbf{ assert } O_t = O'_t = \texttt{i32}
\end{split}
\end{align}
\end{framed}

\subsection{Relations}
\label{sec:rel}

The LLVM statement \texttt{icmp} compares integers or vectors of integers;
and the \texttt{fcmp} statement compares floats of vectors of floats. The
translation produces \texttt{icmp} statements such as:
\begin{verbatim}
  %5 = icmp slt i32 %2, %4
\end{verbatim}

We first define a mapping from B relational operators to LLVM
keywords:
$$D_{op} = \{ = \mapsto \texttt{"eq"}, 
  \ne \mapsto \texttt{"ne"}, 
  < \mapsto \texttt{"slt"}, 
  \le \mapsto \texttt{"sle"}, 
  > \mapsto \texttt{"sgt"}, 
  \ge \mapsto \texttt{"sge"} \}.$$

\begin{framed}
\begin{align}
\begin{split}
  \trad{r, $D$}{rel} \isdef & \textbf{ let } O_p, O_l, O_t =
  \trad{r.left, $D$}{exp} \textbf{ and }
  O'_p, O'_l, O'_t = \trad{r.right, $D$}{exp} \\
  & \quad \textbf{ and } l = \mathcal{V} \textbf{ and } op = D_{op}(\textsf{r.op}) \textbf{ in } \\
  & \textbf{ yields} \\
  & \quad (O_p \conc \\
  & \quad O'_p \conc \\
  & \quad l \conc \texttt{" = icmp "} \conc op \conc \texttt{" "} \conc O_t \conc \texttt{" "} \conc O_l \conc^, O'_l \conc \nl), \\
  & \quad l \\
  & \textbf{ where } \textsf{r.k} = \textsf{BinRel} \\
  & \textbf{ assert } O_t = O'_t = \texttt{i32}
\end{split}
\end{align}
\end{framed}

\subsection{Conditions}
\label{sec:cond}

The translation of conditions takes as input a condition node
\textsf{n}, a dictionary $D$, as well as two labels $L_t$ and $L_f$
that correspond to program locations where the execution shall jump
when the condition is evaluted to true or false, respectively.

The first rule defines the translation of an atomic relation:
\begin{framed}
\begin{align}
\begin{split}
  \trad{}{cond} : & \textsf{ Node} \times dict \times text \times text \rightarrow text \\
  \trad{n, $D$, $L_t$, $L_f$}{cond} \isdef
  & \textbf{ let } O_p, O_l = \trad{n, $D$}{rel} \textbf{ in } \\
  & \textbf{ yields} \\
  & \quad O_p \conc \\
  & \quad \texttt{br i1 } \conc O_l \conc \texttt{" , label \%"} \conc L_t \conc \texttt{", label \%"} \conc L_f \conc \nl  \\
  & \textbf{ where } \textsf{n.k} = \textsf{BinRel}
\end{split}
\end{align}
\end{framed}

Next is the rule for negations:
\begin{framed}
\begin{align}
\begin{split}
  \trad{n, $D$, $L_t$, $L_f$}{cond} \isdef
  & \textbf{ yields} \trad{n.arg, $D$, $L_f$, $L_t$}{cond} \\
  & \textbf{ where } \textsf{n.k} = \textsf{Neg}
\end{split}
\end{align}
\end{framed}

The following rule handles the case of conjunctions. Note that it produces code
with a ``short-cut'' when the evaluation of the first argument yields false:
\begin{framed}
\begin{align}
\begin{split}
  \trad{n, $D$, $L_t$, $L_f$}{cond} \isdef
  & \textbf{ let } L_i = \name \textbf{ in } \\
  & \textbf{ let } O_l = \trad{n.left, $D$, $L_i$, $L_f$}{cond} \textbf{ and } \trad{n.right, $D$, $L_t$, $L_f$}{cond} \textbf{ in } \\
  & \textbf{ yields} \\
  & \quad O_l \conc \\
  & \quad \texttt{; <label>:} \conc L_i \conc \nl \conc \\
  & \quad O_r \\
  & \textbf{ where } \textsf{n.k} = \textsf{Conj}
\end{split}
\end{align}
\end{framed}

The last rule for conditions handles disjunctions. It also produces
code with a short-cut when the first argument evaluates to true:
\begin{framed}
\begin{align}
\begin{split}
  \trad{n, $D$, $L_t$, $L_f$}{cond} \isdef
  & \textbf{ let } L_i = \name \textbf{ in } \\
  & \textbf{ let } O_l = \trad{n.left, $D$, $L_t$, $L_i$}{cond} \textbf{ and } \trad{n.right, $D$, $L_t$, $L_f$}{cond} \textbf{ in } \\
  & \textbf{ yields} \\
  & \quad O_l \conc \\
  & \quad \texttt{; <label>:} \conc L_i \conc \nl \conc \\
  & \quad O_r \\
  & \textbf{ where } \textsf{n.k} = \textsf{Conj}
\end{split}
\end{align}
\end{framed}

\section{Control}
\label{sec:control}

This section addresses how B operations are translated to LLVM
functions. Each operation is translated in turn, and the result of
the translation of individual operations is concatenated to yield the
result of the translation of the operations clause:
\begin{framed}
\begin{align}
  \begin{split}
  \trad{}{ops} : & \textsf{ Node} \rightarrow text \\
  \trad{o}{ops} \isdef 
  & \textbf{ yields } \trad{o}{op} \conc \trad{o.next}{ops}  \\
  & \textbf{ where } \textsf{o} \neq \bot  \textbf{ and } \textsf{o.k} \in \textsf{Op}  \\
  \trad{$\bot$}{ops} \isdef
  & \textbf{ yields } \mty 
  \end{split}
\end{align}
\end{framed}

Three main aspects of the translation are discussed presented, namely
the signature of the operation, information local to the operation,
and the instructions composing the body of the operation. The
translation of the types of local data, as well as that of the
expressions is deferred to sections~\ref{sec:data} and~\ref{sec:expr},
respectively.

One can distinguish three elements in a LLVM function: the signature,
memory allocation for local variables, and blocks of LLVM
statements\footnote{In an attempt to clarify the exposition of our
  ideas, we will call \emph{instructions} the constituents of a B0
  operation, and \emph{statements} those of a LLVM function.}.

To map operations to LLVM functions, one needs to consider the
following:
\begin{itemize}
\item The signature of the function is composed of the name of the
  function, the result type, the name and type of the parameters. The
  signature may carry other information such as linkage, visibility,
  calling convention, etc. The B parser provides this information after
  the type checking phase has been completed.

  As usual in programming languages, LLVM functions may only have one
  result. B0 operations may produce several results. These results are
  also known as output parameters of the operation. The translation of
  output parameters will be function parameters passed by reference,
  i.e. they are pointers. The return type of the LLVM functions
  resulting from the translation of B operations will always be
  \texttt{void}. Translation of operation headers is defined in
  section~\ref{sec:trad-header}

\item The memory to store local variables is allocated on the stack,
  using the statement \verb+alloca+, requiring the type of value that
  need to be stored as well as any memory alignment requirement. The
  Clang compiler places all the stack allocation statements in the
  first block of in the routine. To replicate such behaviour, one
  needs to implement a method to collect the name and type of all the
  local variables used in the operation. Also, the names have to be
  translated to conform to LLVM restrictions. All the data structures
  necessary to implement (straightforwardly) such procedure are
  available in the abstract syntax tree. This part of the translation 
  is defined in section~\ref{sec:trad-alloc}.

\item Finally, the body of the operation has to be transformed to a
  sequence of LLVM blocks of statements. For each kind of instruction
  one (or more) translation function is defined. The details are provided
  in section~\ref{sec:trad-instr}.

\end{itemize}

The translation of an operation is defined by the following rule:
\begin{framed}
\begin{align}
\begin{split}
  \trad{}{op} &: \textsf{ Node} \rightarrow text \\
  \trad{op}{op} \isdef
  & \textbf{ let } O_l = \trad{op}{link} \textbf{ and } O_h, D_h = \trad{op.header}{header} \textbf{ in}  \\
  & \textbf{ let } O_a, D_a = \trad{op.body, $D_h$}{alloc} \textbf{ in}  \\
  & \textbf{ let } O_i = \trad{op.body, $D_a$}{inst} \textbf{ in}  \\
  & \textbf{ yields } \texttt{"declare "} \conc O_l \conc O_h \conc \texttt{"\{"} \conc \nl \conc \\
  & \quad \conc O_a \conc \nl \conc O_i \conc \nl \conc \texttt{"ret void"} \conc \nl \conc \texttt{"\}"} \conc \nl  \\
  & \textbf{ where } \textsf{op} \ne \bot \textbf{ and } \textsf{op.k} = \textsf{Op} 
\end{split}
\end{align}
\end{framed}

The following rule produces the LLVM calling convention keyword
\texttt{internal} that guarantees that the translated B0 local
operations are not visible outside the module:
\begin{framed}
\begin{align}
\begin{split}
  \trad{}{link} &: \textsf{ Node} \rightarrow text \\
  \trad{op}{link} \isdef
  & \textbf{ yields } \texttt{"internal "} \\
  & \textbf{ where } \textsf{op} \ne \bot \textbf{ and } \textsf{op.LocOp} = \top \\
  \trad{op}{link} \isdef
  & \textbf{ yields } \texttt{""} \\
  & \textbf{ where } \textsf{op} \ne \bot \textbf{ and } \textsf{op.LocOp} \ne \top
\end{split}
\end{align}
\end{framed}

\subsection{Sub-routines signatures}
\label{sec:trad-header}

The following rules specify the translation of operation headers. It
is defined through four different patterns, depending on the presence
of input and output parameters. The name of the LLVM function is the
name of the B0 operation prefixed by \mbox{@}. The return type is
always \texttt{void} and the parameter list is made of the translation
of the input and output parameters:

\begin{framed}
  \begin{align}
\begin{split}
    \trad{}{header} &: Hd \rightarrow text \times dict \times dict \\
    \trad{h}{header} \isdef& \textbf{ let } T_n = \textsf{"@"} \conc \textsf{h.ident} 
    \textbf{ and } T_i, D_i = \trad{h.input}{inp} 
    \textbf{ and } T_o, D_o = \trad{h.output}{out} \textbf{ in}  \\
    & \textbf{ let } T_p = T_i \conc^, T_o \textbf{ in}  \\
    & \textbf{ yields } \texttt{"void "} \conc T_n \conc \texttt{"("} \conc T_p \conc
    \texttt{")"}, D_i \cup D_o, \{ \textsf{h.ident} \mapsto T_n \}  \\
    & \textbf{ where } \textsf{h} \ne \bot \textbf{ and } \textsf{h.k} = \textsf{Hd} 
  \end{split}
\end{align}
\end{framed}

Next are the rules that specify recursively the translation of input
parameters. The first pattern is the base case and the second is the
case for parameter lists with two or more elements. The name of the
parameter is the name of the B0 parameter prefixed with \texttt{\%}.
The type of the parameter is translated through specific rules 
defined in section~\ref{sec:data}:

\begin{framed}
  \begin{align}
\begin{split}
    \trad{}{inp} &: \textsf{ Node } \rightarrow text \times dict \\
    \trad{i}{inp} \isdef& \textbf{ let } T_n = \textsf{"\%"} \conc \textsf{i.ident} 
    \textbf{ and } T_t = \trad{i.type}{type} 
    \textbf{ and } T, D = \trad{i.next}{inp} \textbf{ in}  \\
    & \textbf{ yields } T_t \conc \texttt{" "} \conc (T_n \conc^, T), \{ \textsf{i} \mapsto T_n \} \cup D  \\
    & \textbf{ where } \textsf{i} \ne \bot \textbf{ and } \textsf{i.k} = \textsf{Var}  \\
    \trad{$\bot$}{inp} \isdef& \textbf{ yields } \mty, \emptyset 
  \end{split}
\end{align}
\end{framed}

The following rules describe how output parameters are translated. It
is similar to the rules for input parameters. The only difference is
that parameters need to be of pointer type.

\begin{framed}
  \begin{align}
\begin{split}
    \trad{}{out} &: \textsf{ Node } \rightarrow text \times dict \\
    \trad{o}{out} \isdef& \textbf{ let } T_n = \textsf{"\%"} \conc \textsf{o.ident} 
    \textbf{ and } T_t = \trad{o.type}{type} \conc \texttt{"*"}
    \textbf{ and } T, D = \trad{o.next}{out} \textbf{ in}  \\
    & \textbf{ yields } T_t \conc \texttt{" "} \conc (T_n \conc^, T), \{ \textsf{o} \mapsto T_n \} \cup D  \\
    & \textbf{ where } \textsf{o} \ne \bot \textbf{ and } \textsf{o.k} = \textsf{Var}  \\
    \trad{$\bot$}{out} \isdef& \textbf{ yields } \mty, \emptyset  
  \end{split}
\end{align}
\end{framed}

\subsection{Stack allocation}
\label{sec:trad-alloc}

The generation of stack allocation statements processes the parse tree
in a pre-order depth-first traversal, generating one statement for
each local variable declaration found.

\begin{framed}
  \begin{align}
\begin{split}
    \trad{}{alloc} : & \textsf{ Node } \times dict \rightarrow text \times dict \\
    \trad{i, $D$}{alloc} \isdef 
    & \textbf{ let } O_b, D_b = \trad{n, $D$}{alloc}_{base}  \\
    & \textbf{ and } O_n, D_n = \trad{i.next; $D_b$}{alloc}  \\
    & \textbf{ yields } O_b \conc \nl \conc O_n, D_n  \\
    & \textsf{ where } \textsf{i} \ne \bot \textbf{ and } \textsf{i.k} \in \{\textsf{Blk}, \textsf{VarDecl}, \textsf{If}, \textsf{IfBr}, \textsf{While}, \textsf{Case}, \textsf{CaseBr} \}  \\
    \trad{i, $D$}{alloc} \isdef 
    & \textbf{ let } O_n, D_n = \trad{i.next; $D$}{alloc}  \\
    & \textbf{ yields } O_n, D_n  \\
    & \textsf{ where } \textsf{i} \ne \bot \textbf{ and } \textsf{i.k} \not\in \{\textsf{Blk}, \textsf{VarDecl}, \textsf{If}, \textsf{IfBr}, \textsf{While}, \textsf{Case}, \textsf{CaseBr} \}  \\
    \trad{$\bot$, $D$}{alloc} \isdef & \textbf{ yields } \mty, D 
  \end{split}
\end{align}
\end{framed}

One must avoid generating name conflicts in the LLVM output. For each
B0 variable \textsf{v} found, a fresh LLVM name \texttt{l} is thus
created. This association is recorded in the $D$ parameters and results.
\begin{framed}
  \begin{align}
\begin{split}
    \trad{}{alloc}_{base} : & \textsf{ Node} \times dict \rightarrow text \times dict \\
    \trad{i, $D$}{alloc}_{base} \isdef 
    & \textbf{ yields } \trad{i.body, $D$}{alloc}  \\
    & \textbf{ where } \textsf{i} \ne \bot \textbf{ and } \textsf{i.k} \in \{\textsf{Blk}, \textsf{While}, \textsf{IfBr}, \textsf{CaseBr} \}  \\
    \trad{i, $D$}{alloc}_{base} \isdef 
    & \textbf{ yields } \trad{i.branch, $D$}{alloc}  \\
    & \textbf{ where } \textsf{i} \ne \bot \textbf{ and } \textsf{i.k} \in \{\textsf{If}, \textsf{Case} \}  \\
    \trad{i, $D$}{alloc}_{base} \isdef 
    & \textbf{ let } O_l, D_l = \trad{i.ident, $D$}{local} \textbf{ in}  \\
    & \textbf{ let } O_b, D_b = \trad{i.body, $D_l$}{alloc} \textbf{ in}  \\
    & \textbf{ yields } O_l \conc \nl \conc O_b, D_b  \\
    & \textbf{ where } \textsf{i} \ne \bot \textbf{ and } \textsf{i.k} = \textsf{VarDecl} 
  \end{split}
\end{align}
\end{framed}

To illustrate how heap allocation works in LLVM, this is the statement
produced by Clang corresponding to the C declaration
\verb^unsigned short c1^:
\begin{verbatim}
  %c1 = alloca i16
\end{verbatim}
To generate such statement, one needs to create a name
(e.g. \texttt{\%c1}) for the location, and to identify the
corresponding LLVM type (e.g. \texttt{i16}). This is the role of the
last translation function in this section. It produces the desired
source code text and enriches the translation dictionary with the
association between the B0 variable and the created LLVM identifier. It
relies upon an auxiliary translation function $\trad{}{type}$ to be
defined hereafter:
\begin{framed}
\begin{align}
\begin{split}
  \trad{}{local} : & \textsf{ Node } \times dict \rightarrow text \times dict \\
  \trad{v, $D$}{local} \isdef & 
  \textbf{ let } T_v = \fresh(\texttt{"\%"}\conc\textsf{v.ident}, D) \textbf{ in}  \\
  & \textbf{ let } T_n, D_n = \trad{v.next, $D \cup \{ \textsf{v} \mapsto T_v \}$}{local} \textbf{ in}  \\
  & \textbf{ yields } T_v \conc \texttt{" = alloca "} \conc \trad{v.type}{type} \conc \nl \conc T_n, D_n  \\
  & \textbf{ where } \textsf{v} \ne \bot \textbf{ and } \textsf{v.k} = \textsf{Var}  \\
  \trad{$\bot$, $D$}{local} \isdef & 
  \textbf{ yields } \mty, D 
\end{split}
\end{align}
\end{framed}

Following the allocation statement, the LLVM source code contains the
statements that encodes the instructions in the body of the
operation. It is the role of the next section to present how this
translation can be defined.

\subsection{Control flow}
\label{sec:trad-instr}

The translation function $\trad{}{inst}$ shall produce one labeled,
sequential, block of LLVM statements for each of the following in the
B operation: maximal sequences of instructions that do not contain
branching (i.e. conditionals and loops), and evaluation of branching
condition. In the target code, each block has a unique label that
identifies its first statement. The translator has an unbounded supply
of such labels; recall that a new valid LLVM label may be created with
the expression $\name$.

The translation includes parameters to carry additional information
necessary to the translation. One such parameter is a label for the
next instruction to be translated. If its value is $\bot$ then it is
considered missing. Otherwise, it is a valid LLVM label and it shall
be placed in the LLVM output. An auxiliary definition is helpful to
handle the production of labels:
\begin{framed}
\begin{align}
\begin{split}
  \trad{}{inst}_{L} &: Instr \times dict \times text \rightarrow text \\
  \trad{i, $D$, $L$}{inst}_L \isdef 
  & \textbf{yields } L \conc \texttt{:} \trad{i, $D$}{inst}  
\end{split}
\end{align}
\end{framed}

The signature of the translation function for instructions is
\begin{framed}
  \begin{align}
\begin{split}
    \trad{}{inst} &: Instr \times dict \rightarrow text
  \end{split}
\end{align}
\end{framed}
In the forthcoming sections, the definition of this function is given
piecewise, according to the kind of instruction being translated. One
special case is the absence of node representing an instruction:
\begin{framed}
  \begin{align}
\begin{split}
    \trad{$\bot$, $D$}{inst} \isdef \textbf{ yields } \texttt{""}
  \end{split}
\end{align}
\end{framed}

\subsubsection{Block}

Block instructions are structuring constructs that have no operational
intent. The translation is thus defined as the concatenation of the
translation of the block body and the translation of the following
statement.

\begin{framed}
\begin{align}
\begin{split}
  \trad{i, $D$}{inst} \isdef 
  & \textbf{let } O_b = \trad{i.body, $D$}{inst} 
  \textbf{ and } O_n = \trad{i.next, $D$}{inst} \textbf{ in} \\
  & \textbf{yields } O_b \conc O_n  \\
  & \textbf{where } \textsf{i} \ne \bot \textbf{ and } \textsf{i.k} = \textsf{Blk} 
\end{split}
\end{align}
\end{framed}

\subsubsection{If}

To translate an if statement \textsf{i}, each branch is translated in
turn. We assume that each branch has a condition, a body, and a
successor. In the case of an ``else'' branch, the condition is $\bot$.
It is the role of $\trad{}{br}$ to produce the translation for such
branches. Note that at the end of the translation of the body of each
branch, a jump to the instruction following the if statement must be
included. Thus a label $L$ is generated and is a parameter of the
branch translation. It will be used to label the statement
\textsf{i.next}:
\begin{framed}
\begin{align}
\begin{split}
  \trad{i, $D$}{inst} \isdef 
  & \textbf{let } L = \name \textbf{ in} \\
  & \textbf{let } O_b = \trad{i.branch, $D$, $L$}{br} \textbf{ and }
  O_n = \trad{i.next, $D$, $L$}{inst}_L \textbf{ in}  \\
  & \textbf{yields } O_b \conc O_n  \\
  & \textbf{where } \textsf{i} \ne \bot \textbf{ and } \textsf{i.k} = \textsf{If} 
\end{split}
\end{align}
\end{framed}

Now, to translate the branches of an if instruction, one must first
generate code to evaluate the condition of the branch, followed by an
LLVM conditional branching statement translate, and then translate the
body of the branch. In LLVM, a conditional statement is as follows:
\begin{verbatim}
  br i1 %5, label %6, label %10
\end{verbatim}
where \texttt{\%5} is a local variabel of width one bit (since its
type is \texttt{i1}), storing the value of the condition evaluation,
and \texttt{\%6} and \texttt{\%10} are statement labels.

The first rule consider branches that have non-void condition and
sucessor.  Two labels $L_1$ and $L_2$ are created: $L_1$ is the label
of the branch body, and $L_2$ is the label where the code responsible
for the execution of the remaining branches is located. $O_p$ is the
LLVM code responsible for evaluating the condition, $O_l$ being the
LLVM local variable where the result of the evaluation is stored,
$O_b$ and $O_n$ stand for the LLVM code sections responsible for
executing the branch body and the remaining branches, respectively.
Statement block $O_c$ evaluates the condition, where $O_t$ is the name
of the temporary register storing its value in the LLVM:

\begin{framed}
\begin{align}
\begin{split}
  \trad{b, $D$, $L$}{br} : & \textsf{ Node} \times dict \times text \rightarrow text \\
  \trad{b, $D$, $L$}{br} \isdef 
  & \textbf{ let } L_1 = \name \textbf{ and } L_2 = \name \textbf{ in}  \\
  & \textbf{ and } O_p, O_l = \trad{b.cond, $D$, $L_1$, $L_2$}{cond}  \\
  & \textbf{ and } O_b = \trad{b.body, $D$}{inst} \textbf{ and }
  O_n = \trad{b.next, $D$, $L$}{br} \textbf{ in}  \\
  & \textbf{ yields }  \\
  & \quad O_p \conc \nl \conc  \\
  & \quad \texttt{"br i1 "} \conc O_l \conc \texttt{", label \%"} \conc L_1 \conc \texttt{" , label \%"} \conc L_2 \conc \nl \conc  \\
  & \quad L_1 \conc \texttt{:} \conc \nl \conc  \\
  & \quad O_b \conc  \\
  & \quad \texttt{"br "} \conc L \conc \nl \conc  \\
  & \quad L_2 \conc \texttt{":"} \conc \nl \conc  \\
  & \quad O_n  \\
  & \textbf{ where } \textsf{b} \ne \bot \textbf{ and } \textsf{b.k} = \textsf{IfBr} \textbf{ and } \textsf{b.cond} \ne \bot \textbf{ and } \textsf{b.next} \ne \bot 
\end{split}
\end{align}
\end{framed}

The second case is that of the last branch of the if instruction, when
this branch has a condition (i.e. it is an ``elsif''). It is a simplified
version of the previous rule:
\begin{framed}
\begin{align}
\begin{split}
  \trad{b, $D$, $L$}{br} \isdef 
  & \textbf{ let } L_1 = \name \textbf{ and } O_p, O_l, O_t = \trad{b.cond, D}{exp}  \\
  & \textbf{ and } O_b = \trad{b.body, $D$}{inst} \textbf{ in} \\
  & \textbf{ yields }  \\
  & \quad O_p \conc \nl \conc  \\
  & \quad \texttt{"br i1 "} \conc O_l \conc \texttt{", label \%"} \conc L_1 \conc \texttt{" , label \%"} \conc L \conc \nl \conc  \\
  & \quad L_1 \conc \texttt{:} \conc \nl \conc  \\
  & \quad O_b \conc  \\
  & \quad \texttt{"br "} \conc L \conc \nl \conc  \\
  & \textbf{ where } \textsf{b} \ne \bot \textbf{ and } \textsf{b.k} = \textsf{IfBr} \textbf{ and } \textsf{b.next} = \bot  \textbf{ and } \textsf{b.cond} \ne \bot  \\
  & \textbf{ assert } O_t = \texttt{i1} 
\end{split}
\end{align}
\end{framed}

The last case is even simpler; it applies to ``else'' branches:
\begin{framed}
\begin{align}
\begin{split}
  \trad{b, $D$, $L$}{br} \isdef 
  & \textbf{let } O_b = \trad{b.body, $D$}{inst} \textbf{ in} \\
  & \textbf{yields }  \\
  & \quad O_b \conc  \\
  & \quad \texttt{"br "} \conc L \conc \nl \conc  \\
  & \textbf{where } \textsf{b} \ne \bot \textbf{ and } \textsf{b.k} = \textsf{IfBr} \textbf{ and } \textsf{b.next} = \bot  \textbf{ and } \textsf{b.cond} = \bot 
\end{split}
\end{align}
\end{framed}

\subsubsection{Variable declaration}

The static memory allocation has been produced with the $\trad{}{alloc}$
operator. The rest of the translation consists in translating the body
of the variable declaration:

\begin{framed}
\begin{align}
\begin{split}
  \trad{i, $D$}{inst} \isdef 
  & \textbf{ yields } \trad{i.body, $D$}{inst} \\
  & \textbf{ where } \textsf{i} \ne \bot \textbf{ and } \textsf{i.k} = \textsf{VarDecl} 
\end{split}
\end{align}
\end{framed}

\subsubsection{Identity}

If $\textsf{i.k} = \textsf{Id}$, then \textsf{i} has no effect:

\begin{framed}
\begin{align}
\begin{split}
  \trad{i, $D$}{inst} \isdef 
  & \textbf{ yields } \mty \label{trad:id} \\
  & \textbf{ where } \textsf{i} \ne \bot \textbf{ and } \textsf{i.k} \in \{ \textsf{Id}, \textsf{Assert} \} 
\end{split}
\end{align}
\end{framed}

\subsubsection{Becomes equal}

In LLVM, assignment is performed with the \texttt{store} statement, e.g.:
\begin{verbatim}
store i16 %17, i16* %c1
\end{verbatim}
where \texttt{\%17} is the local variable storing the assigned value,
\texttt{\%c1} is the local variable being assigned, and \texttt{i16}
is the type of the assigned value.

The translation generates a sequence of statements $O_l$ to evaluate
the left-hand side in local variable $L_l$ , followed by a sequence of
statements $O_r$ to evaluate the right-hand side in the local variable
$L_r$, followed by a single \verb'store' statement to perform the
assignment:

\begin{framed}
\begin{align}
\begin{split}
  \trad{i, $D$}{inst} \isdef 
  & \textbf{ let } O_l, L_l, T_l = \trad{i.lhs, $D$}{lv} \textbf{ and } O_r, L_r, T_r = \trad{i.rhs, $D$}{exp}  \textbf{ in} \\
  & \textbf{ yields }  \\
  & \quad O_l \conc \nl  \\
  & \quad O_r \conc \nl  \\
  & \quad \texttt{"store "} \conc T_r \conc \texttt{" "} \conc L_r \conc \texttt{", "}
  \conc T_l \conc \texttt{" "} \conc L_l \conc \nl  \\
  & \textbf{ where } \textsf{i} \ne \bot \textbf{ and } \textsf{i.k} = \textsf{BEq}  \\
  & \textbf{ assert } T_l = T_r \conc \texttt{"*"} 
\end{split}
\end{align}
\end{framed}

$\trad{}{lv}$ is responsible for the translation of assignment
targets, also known as \emph{lvalues} in C. The different kinds of
targets are scalar data, array data, array element and field of a
record. Note that, in the LLVM, arrays are so-called first-class
types, and assignment statements may target directly an
array. Considering the restrictions put on the handled B0 elements,
assignment targets shall only be identifiers:

\begin{framed}
\begin{align}
\begin{split}
  \trad{}{lv} &: Node \times dict \rightarrow text \times text \times text \\
  \trad{n, $D$}{lv} \isdef 
  & \textbf{ let } T = \trad{n.type}{type} \textbf{ in }  \\
  & \textbf{ yields } \texttt{""}, D(\textsf{n.ident}), T \conc \texttt{*} \\
  & \textbf{ where } \textsf{n} \ne \bot \textbf{ and } \textsf{n.k} = \textsf{Ident}
\end{split}
\end{align}
\end{framed}

\subsubsection{While}

For the ``while'' instruction the translation creates three statement
labels: $L_1$ is the start of the block that evaluates the loop condition;
$L_2$ is the start of the block that executes the loop body; $L_3$ labels
the exit of the loop, i.e. the next statement. $O_p$ is the
LLVM code responsible for evaluating the condition, $O_l$ being the
LLVM local variable where the result of the evaluation is stored,
$O_b$ and $O_n$ stand for the LLVM code sections responsible for
executing the loop body and the following statements:

\begin{framed}
\begin{align}
\begin{split}
  \trad{i, $D$}{inst} \isdef 
  & \textbf{ let } L_1 = \name \textbf{ and } L_2 = \name \textbf{ and } L_3 = \name \textbf{ in} \\
  & \textbf{ let } O_p, O_l = \trad{i.cond, $D$, $L_2$, $L_3$}{cond} \textbf{ and } O_b = \trad{i.body, $D$}{inst}  \\
  & \textbf{ and } O_n = \trad{i.next, $D$, $L_3$}{inst} \textbf{ in} \\
  & \textbf{ yields }  \\
  & \quad L_1 \conc \texttt{":"} \conc \nl \conc  \\
  & \quad O_p \conc \nl \conc  \\
  & \quad \texttt{"br i1 "} \conc O_l \conc \texttt{", label \%"} \conc L_2 \conc \texttt{" , label \%"} \conc L_3 \conc \nl \conc  \\
  & \quad L_2 \conc \texttt{:} \conc \nl \conc  \\
  & \quad O_b \conc  \\
  & \quad \texttt{"br "} \conc L_1 \conc \nl \conc  \\
  & \quad O_n  \\
  & \textbf{ where } \textsf{i} \ne \bot \textbf{ and } \textsf{i.k} = \textsf{While}  \\
  & \textbf{ assert } O_t = \texttt{"i1"} 
\end{split}
\end{align}
\end{framed}

\subsubsection{Callup}

LLVM has a statement named \texttt{call} for sub-routine calls. For
instance, the following statement has been generated by the Clang
compiler for a call to a C function prototyped as
\verb^void acc(int*, int)^:
\begin{verbatim}
call void @acc(i32* %a, i32 %4)
\end{verbatim}
By convention, each B0 operation are translated to a LLVM with no
return value, and where the input and output parameters of the
operation are translated to parameters passed by value and by
reference, respectively.

Auxiliary definitions $\trad{}{argval}$ and $\trad{}{argref}$ produce
snippets of LLVM code responsible for passing argument by value and
by reference, respectively. The result is a pair of texts, first is
the code for evaluating the arguments, and second is the comma-separated
list of typed arguments:
\begin{framed}
\begin{align}
\begin{split}
  \trad{}{argval} : & \textsf{ Node} \times dict \rightarrow text \times text \times text \\
  \trad{i, $D$}{argval} \isdef 
  & \textbf{ let } O_p, O_v, O_t = \trad{i, D}{exp} \textbf{ and } O'_p, O_r = \trad{i.next, D}{argval} \textbf{ in}  \\
  & \textbf{ yields }  \\
  & \quad O_p \conc O'_p, (O_t \conc \texttt{" "} \conc O_v) \conc^, O_r  \\
  & \textbf{ where } \textsf{i} \ne \bot \textbf{ and } \textsf{i.k} = \textsf{Call}  \\
  \trad{$\bot$, $D$}{argval} \isdef & \texttt{ ""},\texttt{ ""},\texttt{ ""}  \\
  \trad{}{argref} : & \textsf{ Node} \times dict \rightarrow text \times text \times text \\
  \trad{i, $D$}{argref} \isdef 
  & \textbf{ let } O_p, O_v, O_t = \trad{i, D}{exp} \textbf{ and } O'_p, O_r = \trad{i.next, D}{argref} \textbf{ in}  \\
  & \textbf{ yields }  \\
  & \quad O_p \conc O'_p, (O_t \conc \texttt{"* "} \conc O_v) \conc^, O_r  \\
  & \textbf{ where } \textsf{i} \ne \bot \textbf{ and } \textsf{i.k} = \textsf{Call}  \\
  \trad{$\bot$, $D$}{argref} \isdef & \texttt{ ""},\texttt{ ""},\texttt{ ""} 
\end{split}
\end{align}
\end{framed}

The result of these auxiliary routines are combined to form the LLVM code
for the whole function call:

\begin{framed}
\begin{align}
\begin{split}
  \trad{i, $D$}{inst} \isdef 
  & \textbf{ let } O_p, O_v = \trad{i.in, D}{argval} \textbf{ and } O'_p, O_r = \trad{i.ou, D}{argref} \textbf{ in} \\
  & \textbf{ yields }  \\
  & \quad O_p \conc O'_p \conc  \\
  & \quad \texttt{"call void "} \conc \trad{i.op, $D$}{exp} \conc \texttt{"("} \conc O_v \conc^, O_r \conc \texttt{")"} \nl  \\
  & \textbf{ where } \textsf{i} \ne \bot \textbf{ and } \textsf{i.k} = \textsf{Call} 
\end{split}
\end{align}
\end{framed}

\subsubsection{Case}

This instruction shall be translated to a \verb'switch' statement.
For the sake of simplicity, we assume that case statement without
else branch are changed so that there is such a branch, with a single
identity instruction.

The LLVM switch statement is basically a jump table with the following
concrete syntax (example):
\begin{verbatim}
switch i32 %val, label %otherwise [ i32 0, label %onzero
                                    i32 1, label %onone
                                    i32 2, label %ontwo ]
\end{verbatim}

The translation of a case statement shall thus produce a labeled block
for each branch, and a jump for each value in each branch. We first
provide an auxiliary definition that yields a mapping from branches to
labels.
\begin{framed}
  \begin{align}
\begin{split}
    \trad{}{lab} : & \textsf{ Node} \rightarrow \textsf{Node} \rightarrow text \\
    \trad{br}{lab} \isdef 
    & \textbf{ yields } \{ \textsf{br} \mapsto  \name \} \cup
    \trad{br.next}{lab}  \\
    & \textbf{ where } \textsf{br} \ne \bot \textbf{ and } \textsf{br.k} = \textsf{CaseBr}  \\
    \trad{$\bot$}{lab} \isdef & \textbf{ yields } \{\} 
  \end{split}
\end{align}
\end{framed}

Next we provide definitions that yield the jump table for all
but the else branch. First, $\trad{}{jump}$ takes as input a
case branch node, and the mapping to labels and recurses over
all branches and branch values:
\begin{framed}
  \begin{align}
\begin{split}
    \trad{}{jump} : & \textsf{ Node} \times (\textsf{Node} \rightarrow text) \rightarrow text \\
    \trad{br, $lab$}{jump} \isdef
    & \textbf{ yields }  \\
    & \quad \trad{br.val, $lab($br$)$}{jump}_v \conc  \\
    & \quad \trad{br.next, $lab$}{jump}  \\
    & \textbf{ where } \textsf{br} \ne \bot \textbf{ and } \textsf{br.k} = \textsf{CaseBr}  \\
    \trad{$\bot$, $lab$}{jump} \isdef
    & \textbf{ yields } \texttt{""} 
  \end{split}
\end{align}
\end{framed}

$\trad{}{jump}_v$ is applies to branch values, and has as second argument a single label $l$:
\begin{framed}
  \begin{align}
\begin{split}
    \trad{}{jump}_v : & \textsf{ Node} \times text \rightarrow text \\
    \trad{val, $l$}{jump}_v \isdef
    & \textbf{ yields }  \\
    & \quad \trad{val.type}{type} \conc \texttt{" "} \conc 
    \trad{val.value}{value} \conc \texttt{", label \%"} \conc l \conc \nl \conc \\
    & \quad \trad{val.next, $l$}{jump}  \\
    & \textbf{ where } \textsf{val} \ne \bot \textbf{ and } \textsf{val.k} = \textsf{Value}  \\
    \trad{$\bot$, $l$}{jump}_v \isdef
    & \textbf{ yields } \texttt{""} 
  \end{split}
\end{align}
\end{framed}

The third and last auxiliary definition $\trad{}{blocks}$ is
responsible for the generation of the statement blocks that correspond
to the branches of a case statement. The parameters are the branch
itself, the variable dictionnary $D$, the branch labeling function
$lab$ and a label $l_f$ where the execution must proceed after
completing the branch:
\begin{framed}
\begin{align}
\begin{split}
  \trad{br, $D$, $lab$, $l_f$}{block} \isdef 
  & \textbf{yields } \\
  & \quad lab(\textsf{br}) \conc \texttt{":"} \conc \nl \conc  \\
  & \quad \trad{br.body, $D$}{inst} \conc  \\
  & \quad \texttt{"br "} \conc l_f \conc \nl \conc  \\
  & \quad \trad{br.next, $D$, $lab$, $l_f$}{block}  \\
  & \textbf{where } \textsf{br} \ne \bot \textbf{ and } \textsf{br.k} = \textsf{CaseBr}  \\
  \trad{$\bot$, $D$, $lab$, $l_f$}{block} \isdef 
  & \textbf{yields } \texttt{""}
\end{split}
\end{align}
\end{framed}

Finally, the following definition details the translation of the case
instruction. First, the tested expression is evaluated, then a switch
statement is emitted, including the jump table, followed by the
translation of the different branches and ended with a label where all
branches eventually jump to.

\begin{framed}
\begin{align}
\begin{split}
  \trad{i, $D$}{inst} \isdef 
  & \textbf{let } lab = \trad{i.branches}{lab} \textbf{ and } p_e, l_e, t_e = \trad{i.exp, D}{exp} \textbf{ in} \\
  & \textbf{let } l_d, l_f = \name, \name \textbf{ and } 
  jumps = \trad{i.branches, $lab$}{jump} \textbf{ and }  \\
  & \quad blocks = \trad{i.branches, $D$, $lab$, $l_f$}{block} \textbf{ in}  \\
  & \textbf{yields }  \\
  & \quad p_e \conc  \\
  & \quad \texttt{"switch "} \conc t_e \conc \texttt{" "} \conc l_e \conc 
  \texttt{ default } \conc l_d \conc \texttt{"["} \conc \nl  \\
  & \quad \quad jumps \conc \texttt{"]"} \conc \nl  \\
  & \quad blocks \conc  \\
  & \quad l_f \conc \texttt{:} \conc \nl  \\
  & \textbf{where } \textsf{i} \ne \bot \textbf{ and } \textsf{i.k} = \textsf{Case}
\end{split}
\end{align}
\end{framed}

As in the case of the if statement, the translation of the following 
instruction initiates a new block, labeled with say \textit{next}. 
Each branch of the case is translated to a block that ends with
an unconditional jump \verb'br label' $\textit{next}$.

% Case_instruction ::=
% ``CASE'' Simple_term ``OF''
% ``EITHER'' Simple_term +'','' ``THEN'' Instruction ( ``OR'' Simple_term +'','' ``THEN'' Instruction )* [ ``ELSE'' Instruction ]
% ``END''
% ``END''

% The CASE substitution is used to define for a program, various
% possible behaviors depending on the value of an expression. Each
% EITHER and OR branch is made up of a non empty list of constants. The
% values of constants in the set of branches must be distinct two by
% two. If the value of the expression belongs to one of the branches,
% then the substitution of this branch is executed. If not, the
% substitution in the ELSE branch is applied, if this latter branch is
% absent, it will by default perform the identity substitution. The
% behavior of this substitution is therefore deterministic and always
% feasible.

%  ; Implement a jump table:
%  switch i32 %val, label %otherwise [ i32 0, label %onzero
%                                      i32 1, label %onone
%                                      i32 2, label %ontwo ]

\subsubsection{Assert}

This instruction has no operational meaning and it needs not be
translated (it might have some effect on the control flow graph, or
may be translated for debugging purposes --- see debugging
intrinsics). Rule~\ref{trad:id} handles this case.

\section{Modularity}

\subsection{The SEES clause}

Suppose a machine $M$ is seen from the implementation being translated
to LLVM. For each constant, set, variable, and operation found in $M$,
the LLVM code must include a corresponding declaration. As an
illustration, consider the following C code:
\begin{verbatim}
extern int f(int, int, int *);
extern char const k1;
extern int k2;
\end{verbatim}
The corresponding LLVM produced by the C compiler is:
\begin{verbatim}
declare i32 @f(i32, i32, i32*)
@k1 = external constant i8
@k2 = external global i32
\end{verbatim}
Also, for the following C++ code:
\begin{verbatim}
class t1;
extern t1 c1;
\end{verbatim}
the compiler produces the following LLVM:
\begin{verbatim}
%class.t1 = type opaque
@c1 = external global %class.t1
\end{verbatim}

\subsection{The PROMOTES clause}

For each promoted operation, a sub-routine definition is generated.
The definition is just a call to the sub-routine corresponding to
the operation being promoted. As an inspiration, let us consider
the following C code:
\begin{verbatim}
extern int f(int);

int promote_f(int p)
{
  return f(p);
}
\end{verbatim}
The compiler produces the following LLVM code:
\begin{verbatim}
define i32 @promote_f(i32 %p) nounwind uwtable ssp {
  %1 = tail call i32 @f(i32 %p) nounwind
  ret i32 %1
}
declare i32 @f(i32)
\end{verbatim}
Note that the function declaration found in the last line shall be
emitted during the processing of the IMPORTS clause for the machine
containing the promoted operation.

\texttt{nounwind}: the called function never returns an unwind or exceptional
control flow.

\section{Completing the definition}

This section is used to document a list of tasks to complete the
definition of LLVM code generation from B0. It is based on the
directory of tests for the C code generation tool ComenC.

The ComenC tests are structured in directories. Each directory tests
the support for some aspect of B0. For each such directory, the
document summarizes the content of each test file and discuss how our
LLVM code generation from B0 needs to be updated.

\subsection{Directory OPERLOC}

This directory is used to test and validate support for ``local operations''.

Local operations are operations that are specified and implemented in
the implementation: they do not refine an operation of the refined
component, they are just a code structuring (factoring) construct. The
specification is located in the "local\_operations" clause, and the
implementation in the "operations clause. Local operations are only
visible to the operations of the machine, and the generated LLVM code
must include corresponding visibility directives.

\subsubsection{Benchmark description}

\begin{description}

\item[operloc\_comments] The implementation has
local operations with comments.

\item[operloc\_empl\_oploc\_ds\_spec] The implementation has three bare
no-op local operations

\item[operloc\_emplacement1] The implementation has three local
operations called in a refined operation.

\item[operloc\_emplacement2] Implementation has three local operations,
two refined operations, two local operations calling the third one.

\item[operloc\_import] Implementation has one local operation, one
imported machine with one operation, the local operations called the
operation of the imported machine

\item[operloc\_ordre1, operloc\_ordre2] Local operations calls other local
operation(s) sometimes appearing further.

\item[operloc\_params] Implementation has local operations with input
and output parameters, and accessing a visible variable

\item[operloc\_promues] The implementation refines an operation by
  promoting an operation of an imported machine, the local operation
  calls the promoted operation. The definition of the local operation
  has a local variable and writes a concrete variable.

\end{description}

\subsubsection{Actions to handle benchmarks}

\begin{enumerate}
\item Emit the keyword \texttt{internal} in the signature of
  operations that appear in the ``local\_operations'' clause.
\item LLVM does not require declaration of sub-routines before their
  use, so code generation shall not worry about ordering operation
  definitions or emitting declarations.
\item Considering promoted operations, it seems that declaring their
  signature is sufficient. For the following C declaration:
\begin{verbatim}
  extern void promop(void);
\end{verbatim}
  the C compiler produces:
\begin{verbatim}
  declare void @promop()
\end{verbatim}
\item Optionally, check which local operations are actually (directly
  or indirectly) referenced by refined operations; only emit code for
  those operations that are used.
\end{enumerate}

\subsection{Directory ORDREDECL}

This directory contains a single benchmark to test modularity constructs:
\begin{description}
\item[test\_ordre] Contains a development with a specification and an
  implementation. The implementation contains imports, extends and
  promotes clauses. Also the benchmark contains declarations of sets
  (abstract and enumerated) and concrete constants (integers and
  intervals).
\end{description}

For each abstract set, a LLVM type is generated. Eventually, if this
abstract set is valued in a B development it will be an integer range.
Since the size of this range remains undefined, there are two
possibilities:

(i) either choose the largest possible integer type available
in the LLVM platform,

(ii) or define it as an opaque type in LLVM (however those seem to be
usually employed for forward declared structure in C or classes
declarations in C++).

\section{Next steps}

Follows a list of tasks to be executed, in no particular order,
and possibly concurrently, to complete this work:
\begin{itemize}

\item Specify code generation for a larger class of data types: arrays,
records, integer intervals, enumerations.

\item Specify modular code generation: sees, imports, extends...

\item Case studies: check that LLVM code is processable by existing
compilation tools.

\item Verify if parse tree produced by BCompiler is close enough to 
  the hypothetic parse tree used in the document. Possibly make adjustements
  to rules to this reality.

\item Validate ideas with manual production of LLVM source files from
simple B0 implementations.

\item Implement.

\end{itemize}

\end{document}

\begin{align}
IMPLEMENTATION & ::= "IMPLEMENTATION" Header \\
& Clause_refines \\
& Clause_implementation* \\
& "END"
\end{align}

Clause_implementation ::= Clause_sees
| Clause_imports
| Clause_promotes
| Clause_extends_B0
| Clause_sets
| Clause_concrete_constants
| Clause_properties
| Clause_values
| Clause_concrete_variables
| Clause_invariant
| Clause_assertions
| Clause_initialization_B0
| Clause_operations_B0