### 2013-09-23 translation.tex
Profound changes in the text laying out the principles for composing translation of different modules related by import clauses. The specification of the translation has not been updated accordingly and is deprecated.

\documentclass{llncs}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amssymb}
\usepackage[fleqn]{amsmath}
\usepackage[colorlinks=false]{hyperref}
\usepackage{xspace,framed,here,keystroke}
\usepackage{graphicx}
\pagestyle{headings}
\title{Translation of B Implementations to the LLVM: Control and Modularity}
\author{David Déharbe \and Valério Medeiros Jr} 
\institute{Program of Graduate Studies in Systems and Computing (PPgSC) \\
  Federal University of Rio Grande do Norte \\
  Natal (Brazil)}
\date{May 2013}

\newcommand{\trad}[2]{\ensuremath{\lVert \textsf{#1} \rVert^{\textit{#2}}}}
\newcommand{\nl}[0]{\ensuremath{\downarrow}}
\newcommand{\mty}[0]{\texttt{""}}
\DeclareMathOperator{\conc}{\diamond}
\DeclareMathOperator{\isdef}{\equiv}
\DeclareMathOperator{\dom}{\mbox{dom}}
\DeclareMathOperator{\lbl}{\mathcal{L}()}
\DeclareMathOperator{\variable}{\mathcal{V}()}
\newcommand{\llvm}[1]{\texttt{#1}}
\newcommand{\B}[1]{\textsf{#1}}
\newcommand{\lalt}[0]{$\langle$\xspace}
\newcommand{\ralt}[0]{$\rangle$\xspace}
\newcommand{\alt}[0]{$\mid\,$}
\newcommand{\ListOf}[1]{$\mbox{#1}^+$}
\newcommand{\nt}[1]{{\normalfont\textit{#1}}}
\newcommand{\Dict}[0]{\mathbb{D}}
\newcommand{\Text}[0]{\mathbb{T}}
\newcommand{\IF}[0]{\textbf{ if }}
\newcommand{\ELSIF}[0]{\textbf{ else if }}
\newcommand{\ELSE}[0]{\textbf{ else }}
\newcommand{\THEN}[0]{\textbf{ then }}
\newcommand{\LET}[0]{\textbf{ let }}
\newcommand{\IN}[0]{\textbf{ in }}
\newcommand{\AND}[0]{\textbf{ and }}
\newcommand{\PH}[1]{\framebox{$#1$}}
\newcommand{\sep}[0]{\otimes}
\newcommand{\intf}[0]{\ensuremath{\mathbb{I}}}
\newcommand{\Global}[0]{\ensuremath{\sf\Gamma}}
\newcommand{\local}[0]{\ensuremath{\sf\lambda}}
\newcommand{\opmap}[0]{\ensuremath{\sf\Omega}}
\newcommand{\idx}[0]{\ensuremath{\sf\Pi}}
\newcommand{\state}[0]{\ensuremath{\sf\Theta}}
\newcommand{\self}[0]{\ensuremath{\sf\Sigma}}
\newcommand{\init}[0]{\ensuremath{\mathsf{I}}}
\newcommand{\tradi}[2]{\ensuremath{\langle \textsf{#1} \rangle^{\textit{#2}}}}

\newcommand{\DD}[1]{\marginpar{\tiny{DD: #1}}}

\begin{document}
\maketitle

\begin{abstract}
  The aim of this work is to lay the foundations of a multi-platform code
  generator for the B method. In particular, this paper presents a translation
  procedure from a large subset of the B language for implementations towards
  LLVM source code. This translation is defined formally as a set of rules
  defined recursively on the abstract syntax for B implementations. It handles
  the following elements of the B language: simple data types, imperative
  instructions and component compositions. 
\end{abstract}

\section{Introduction}

B is a formal refinement-based software design method~\cite{Abrial1996}. It has
a single language encompassing abstract constructs suitable for specification,
as well as classic imperative constructs for computer programming. The starting
point of a B development is a specification, called a \emph{machine}, that is
incrementally refined to an \emph{implementation}, where only imperative-like
constructs may be employed~\cite{Clearsy}. Such implementation may then be
translated to source code in a programming language, say C~\cite{ComenC} or
Ada. All steps in the B method are formally verified using certified theorem
proving technologies. However the translation to some programming language and
its subsequent compilation to the target platform do not benefit from the same
mathematical rigor. In practice, redundancy in the tool chains and execution
platform are employed to increase the level of confidence to the desired levels.

The goal of this work is to connect the B method to the LLVM compilation
framework~\cite{Lattner04LLVM}.  LLVM is a free, open-source project upon which
many compiling contemporary technologies and tools are based. LLVM provides an
intermediate assembly language upon which may be applied techniques such as
optimization, static analysis, code generation, debugging, etc. Providing a
connection from the B method to this framework gives the users of the B method
access to these technologies. Our approach is to define a translation from B0
(the subset of the B language that is used to describe imperative programs) to
the LLVM intermediate format.

Considering the correctness of the proposed translation, we shall rely upon
human inspection for the time being. The standard uniform semantic framework for
B and LLVM that would be necessary to carry out a proof of this correctness does
not exist (yet). A possible approach to mitigate the risk of having an error in
the translation would be to include assertions in the target code. Indeed the
artifacts produced in the B development have many such assertions and it would
be possible to translate at least some of them (B has an assertion instruction
that introduces additional verification conditions in the development). The
produced code could then be (at least partially) verified by testing the absence
of assertion violations.

\paragraph{Overview.} Following this introduction, the paper is organized as
follows. The implementation constructs of the B language, and selected elements
of the LLVM are presented in section~\ref{sec:b-impl}
and~\ref{sec:llvm}. Section~\ref{sec:overview} presents a bird's-eye view that
is then detailed in sections~\ref{sec:module} (for component modularity)
~\ref{sec:data} (for data aspects),~\ref{sec:expr} (expressions) and
instructions~\ref{sec:control}. We conclude and discuss future work
in~\ref{sec:conclusion}.

\section{Implementation Language in the B-method}
\label{sec:b-impl}

This section describes the B implementation constructs that are considered in
the compilation to the LLVM. This description introduces the notations used in
the definition of the translation. The elements of the B language are denoted
using a \B{sans serif} font.

Figure~\ref{tab:node-attr} summarizes the structure
of the abstract syntax that is relevant in the definition of the proposed
translation and introduces identifiers used in the definition of the
translation. For instance, and for the purpose of the translation, the relevant
components of a B implementation are: \B{id}, the name identifying the
implementation within the project; \B{see}, a set of references to other modules
containing definitions used in the implementation; \B{const}, a set of constant
names, their type and their values; \B{impo}, a sequence of optionally
prefixed B machines corresponding to external modules instantiated in the
implementation; \B{var}, a set of variable names, their type and possibly
additional functional restrictions forming the rest of the state; \B{init}, a
sequence of instructions executed upon initialization of the implementation;
\B{op}, a set of operations, each being an algorithmic description of the
different functionalities provided by the implementation.\DD{Include the
possibility that the import is a machine without implementation. Instead it
has a list of typed, implementable, variables.}

The initialization and the operations are defined as (sequences of)
instructions. Operations also have a name, inputs and outputs. Possible
instructions are the classic imperative constructs: variable assignment, if and
case conditional, while loop, block with or without local variables, and
operation calls. The if constructs may have several branches, each with a
condition as well as an optional else branch. Called operations may be either
local operations or operations of instances of external modules, in this case
they have an optional prefix identifying the module.

Note that the translation presented in this paper only considers types \B{INT}
(integer) and \B{BOOL} (Boolean). Accordingly, the expression language consists
in arithmetic operations \B{-} (unary and binary), \B{+}, \B{*}, \B{/}, \B{mod},
% \B{**} (exponentiation), 
\B{succ} and \B{pred} and Boolean operations
\B{$\land$}, \B{$\lor$}, \B{$\neg$} as well as relations \B{=}, \B{$\neq$},
\B{$>$}, \B{$<$}, \B{$\leq$} and \B{$\neq$}. Atomic expressions are identifiers,
Boolean constants \B{FALSE} and \B{TRUE}, integer literals, and integer
constants \B{MAXINT} and \B{MININT}. The development through the application of
the B method ensures the absence of overflows and underflows.

\begin{figure}[H]
  \begin{center}
    {\footnotesize
      \frame{
    \begin{tabular}[t]{ccccc}
      \begin{tabular}[t]{rcl}
        \hline
        \multicolumn{3}{|c|}{Implementation: \B{Impl}} \\
        \hline
        \B{id} & : & \B{Name} \\
        \B{impo} & : & seq \B{Module} \\
        \B{const} & : & seq \B{Cons} \\
        \B{var} & : & seq \B{Vari} \\
        \B{init} & : & seq \B{Inst} \\
        \B{op} & : & seq \B{Oper} 
        \\
        \hline
        \multicolumn{3}{|c|}{Modules: \B{Module}} \\
        \hline
        \B{mach} & : & \B{MachRef} \\
        \B{pre} & : & opt \B{Name} 
        \\
        \hline
        \multicolumn{3}{|c|}{Machine: \B{MachRef}} \\
        \hline
        \B{spec} & : & \B{MachInterface} \\
        \B{impl} & : & opt \B{Impl} \\
        \hline
        \multicolumn{3}{|c|}{Machine interface: \B{MachInterface}} \\
        \hline
        \B{id} & : & \B{Name} \\
        \B{const} & : & seq \B{Cons} \\
        \B{var} & : & seq \B{Vari} \\
        \B{init} & : & seq \B{Inst} \\
        \B{op} & : & seq \B{Oper} 
        \\
        \hline
        \multicolumn{3}{|c|}{Constant: \B{Cons}} \\
        \hline
        \B{id} & : & \B{Name} \\
        \B{type} & : & \B{Type} \\
        \B{val} & : & \B{Value} \\
        \hline
        \multicolumn{3}{|c|}{Variable: \B{Vari}} \\
        \hline
        \B{id} & : & \B{Name} \\
        \B{type} & : & \B{Type} \\
        \B{scope} & : & \B{Impl} \alt \B{Oper}
        \\
      \end{tabular}
      & &
      \begin{tabular}[t]{rcl}
        \hline
        \multicolumn{3}{|c|}{Operation: \B{Oper}} \\
        \hline
        \B{id} & : & \B{Name} \\
        \B{inp} & : & seq \B{Vari} \\
        \B{out} & : & seq \B{Vari} \\
        \B{body} & : & \B{Inst}
        \\
        \hline
        \multicolumn{3}{|c|}{Instruction: \B{Inst}} \\
        \hline
        \multicolumn{3}{c}{\B{Blk} \alt \B{VarD} \alt \B{If} \alt \B{BEq} \alt } \\
        \multicolumn{3}{c}{\B{Call} \alt \B{While} \alt \B{Case} \alt \B{Skip}} \\
        \hline
        \multicolumn{3}{|c|}{Block: \B{Blk}} \\
        \hline
        \B{body} & : & seq \B{Inst} \\
        \hline
        \multicolumn{3}{|c|}{Variable declaration: \B{VarD}} \\
        \hline
        \B{vars} & : & seq \B{Name} \\
        \B{body} & : & seq \B{Inst} \\
        \hline
        \multicolumn{3}{|c|}{If: \B{If}} \\
        \hline
        \B{branches} & : & seq \B{IfBr} \\
        \hline
        \multicolumn{3}{|c|}{Becomes equal: \B{Beq}} \\
        \hline
        \B{lhs} & : & \B{Vari} \\
        \B{rhs} & : & \B{Expr} \\
        \hline
        \multicolumn{3}{|c|}{Operation call: \B{Call}} \\
        \hline
        \B{op} & : & \B{Oper} \\
        \B{inp} & : & seq \B{Expr} \\
        \B{out} & : & seq \B{Name} \\
        \B{inst} & : & opt \B{Impo}
        \\
        \hline
        \multicolumn{3}{|c|}{Case conditional: \B{Case}} \\
        \hline
        \B{expr} & : & \B{Expr} \\
        \B{branches} & : & seq \B{CaseBr}
      \end{tabular}
      & &
      \begin{tabular}[t]{rcl}
        \hline
        \multicolumn{3}{|c|}{Case branch: \B{CaseBr}} \\
        \hline
        \B{val} & : & seq \B{Value} \\
        \B{body} & : & \B{Inst}
        \\
        \hline
        \multicolumn{3}{|c|}{While loop: \B{While}} \\
        \hline
        \B{cond} & : & \B{Pred} \\
        \B{body} & : & \B{Inst}
        \\
        \hline
        \multicolumn{3}{|c|}{If branch: \B{IfBr}} \\
        \hline
        \B{cond} & : & opt \B{Pred} \\
        \B{body} & : & \B{Inst} \\
        \hline
        \multicolumn{3}{|c|}{Expression: \B{Expr}} \\
        \hline
        \multicolumn{3}{c} {\B{Lit} \alt \B{Term} \alt \B{Pred} \alt \B{Vari}} \\
        \hline
        \multicolumn{3}{|c|}{Term expression: \B{Term}} \\
        \hline
        \B{op} & : & \B{ArithOp} \\
        \B{args} & : & seq \B{Exp} \\
        \hline
        \multicolumn{3}{|c|}{Predicate: \B{Pred}} \\
        \hline
        \multicolumn{3}{c}{\B{Form} \alt \B{Comp}} \\
        \hline
        \multicolumn{3}{|c|}{Boolean formula: \B{Form}} \\
        \hline
        \B{op} & : & \B{BoolOp} \\
        \B{args} & : & seq \B{Pred} \\
        \hline
        \multicolumn{3}{|c|}{Comparison: \B{Comp}} \\
        \hline
        \B{op} & : & \B{RelOp} \\
        \B{arg1}, \B{arg2} & : & \B{Exp} \\
        \hline
        \multicolumn{3}{|c|}{Data type: \B{Type}} \\
        \hline
        \multicolumn{3}{c}{\B{INT} \alt \B{BOOL}}
      \end{tabular}
    \end{tabular}
  }
  }
    \caption{Abstract syntax structure of a B implementation.  For each abstract
      syntax element, we give the name of the class (e.g. \B{CaseBr} for a
      branch in a case instruction, and the different attributes of the class),
      or a list of the possible sub-classes (e.g. \B{Inst} for the different
      kinds of instructions).
      Sequence and optional attributes are denoted ``seq'' and ``opt'',
      respectively}
    \label{tab:node-attr}
  \end{center}
\end{figure}

\section{Target LLVM Subset}
\label{sec:llvm}

LLVM aims to be a general, language agnostic, intermediate representation for
compilers. The compiler front-end translates the source programming language to
LLVM source code, upon which optimization and other static analysis tasks may be
performed, and the compiler back-end translates to the target platform assembly
language. One key feature of LLVM is that it is a single-static assignment (SSA)
language: a variable may only be assigned once. So for each assignment to a
variable found in the source language, the front-end shall generate a new
temporary variable in the LLVM program. For illustration,
figure~\ref{fig:ex-llvm} presents a LLVM program (for clarity, the corresponding
C code appears on the left) with two temporary variables \llvm{\%0} and
\llvm{\%1}.
\begin{figure}
\vspace*{-3em}
  \begin{minipage}[t]{.3\textwidth}
    \begin{tabbing}
      foo \= \kill  \\
      void inc(int * pi) \{\\
      \> *pi += 1; \\
      \}\end{tabbing}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{.6\textwidth}
    \begin{tabbing}
      foo \= \kill  \\
      \llvm{define void @inc(i32* \%pi) nounwind \{} \\
      \llvm{entry:}                      \\
      \> \llvm{\%0 = load i32* \%pi}     \\
      \> \llvm{\%1 = add i32 \%0, 1}     \\
      \> \llvm{store i32 \%1, i32* \%pi} \\
      \> \llvm{ret void} \\
      \llvm{\}}
    \end{tabbing}
  \end{minipage}
  \caption{Simple example of a C function and its corresponding LLVM function.
    The first line contains the signature: return type \llvm{void}, the name
    \llvm{@inc} and one parameter named \llvm{\%pi} and typed \llvm{i32*}.  Next
    is the body with a single block, labeled \llvm{entry}, and temporary
    variables \llvm{\%0} and \llvm{\%1}, created in the conversion to SSA. The
    block has four instructions: \llvm{load}, \llvm{add}, \llvm{store} and
    \llvm{ret}. For instance, \llvm{\%1 = add i32 \%0, 1} is an addition
    (\llvm{add}), has result type \llvm{i32} and assigns to
    \llvm{\%1} the sum of variable \llvm{\%0} and integer
    literal \llvm{1}.}
  \label{fig:ex-llvm}
\end{figure}

The introduction to the LLVM that follows is focused in the features
that are relevant for the representation of B
implementations. Figure~\ref{fig:llvm-grammar} presents a grammar of this subset
of the LLVM.

\begin{figure}
  \begin{center}
    \begin{tabular}{rcl}
      \nt{module} & ::= & \ListOf{\nt{item}} \\
      \nt{item} & ::= & \nt{const\_decl} \alt \nt{function\_decl} \alt \nt{type\_def} 
      \alt \nt{const\_def} \alt \nt{var\_def} \alt \nt{function\_def} \\
      \nt{const\_decl} & ::= & \nt{name} \llvm{=} \llvm{external} \llvm{constant} \nt{type} \\
      \nt{type\_def} & ::= & \nt{name} \llvm{=} \llvm{type} \nt{type} \\
      \nt{type} & ::= & \llvm{void} \alt \nt{itype} \alt \llvm{\{} \ListOf{\nt{type}} \llvm{\}} \alt \nt{type}\llvm{*} \\
      \nt{const\_def} & ::= & \nt{name} \llvm{=} \llvm{constant} \nt{type} \nt{iliteral} \\
      \nt{var\_def} & ::= & \nt{name} \llvm{=} \llvm{common} \llvm{global} \nt{type} \llvm{zeroinitializer} \\
      \nt{function\_decl} & ::= & \llvm{declare} \nt{type} \nt{name} \llvm{(} \ListOf{\nt{type}} \llvm{)}\\
      \nt{function\_def} & ::= & \llvm{define} \nt{type} \nt{name} \llvm{(} \ListOf{\nt{param}} \llvm{)} \llvm{\{} \ListOf{\nt{block}} \llvm{\}} \\
      \nt{param} & ::= & \nt{type} \nt{name} \\
      \nt{block} & ::= & \nt{lbl} \llvm{:} \ListOf{\nt{inst}} \\
      \nt{inst} & ::=  & \nt{name} \llvm{=} \llvm{alloca} \nt{type} \\
      & \alt & \nt{name} \llvm{=} \lalt \llvm{add} \alt \llvm{sub} \alt \llvm{mul} \alt \llvm{sdiv} \alt \llvm{srem} \ralt \nt{itype} \nt{exp} \llvm{,} \nt{exp} \\
      & \alt & \nt{name} \llvm{=} \llvm{icmp} \lalt \llvm{eq} \alt \llvm{ne} \alt \llvm{sgt} \alt \llvm{sge} \alt \llvm{slt} \alt \llvm{sle} \ralt \llvm{i1} \nt{exp} \llvm{,} \nt{exp}\\
      & \alt & \nt{name} \llvm{=} \llvm{call} \nt{type} \llvm{(} \ListOf{\nt{arg}} \llvm{)} \\
      & \alt & \nt{name} \llvm{=} \llvm{getelementptr} \nt{type} \llvm{*} \nt{exp}\llvm{,} \nt{index}\llvm{,} \nt{index} \\
      & \alt & \nt{name} \llvm{=} \llvm{load} \nt{type} \nt{exp} \\
      & \alt & \llvm{store} \nt{type} \nt{exp}, \nt{type} \llvm{*} \nt{exp} \\
      & \alt & \llvm{br} \llvm{i1} \nt{exp} \llvm{,} \llvm{label} \nt{lbl} \llvm{,} \llvm{label} \nt{lbl} \\
      & \alt & \llvm{br} \llvm{label} \nt{lbl} \\
%       & \alt & \llvm{switch} \nt{type} \nt{exp} \llvm{,} \llvm{branch} \nt{lbl} \llvm{[} \ListOf{\nt{branch}} \llvm{]} \\
      & \alt & \llvm{ret} \lalt \nt{type} \nt{exp} \alt \llvm{void} \ralt \\
      \nt{exp} & ::= & \nt{name} \alt \nt{iliteral} \alt \llvm{getelementptr} \llvm{(} \nt{type} \nt{exp} \llvm{,} \nt{index} \llvm{,} \nt{index} \llvm{)} \\
      \nt{index} & ::= & \nt{itype} \nt{iliteral} \\
      \nt{branch} & ::= & \nt{iliteral} \nt{iliteral} \nt{lbl} \\
      \nt{arg} & ::= & \nt{type} \nt{exp}
    \end{tabular}
  \end{center}
  \caption{Grammar of the target LLVM subset: \nt{itype}, \nt{iliteral}, \nt{lbl}
    and \nt{name} correspond respectively to integer types, integer literals,
    labels and names. Choices are separated by \alt and optionally delimited by
    \lalt and \ralt.  The \ListOf{} superscript denotes a comma-separated list of
    elements of the annotated entity.}
  \label{fig:llvm-grammar}
\end{figure}

LLVM programs are organized into modules, one per translation unit. A module may
contain declarations of external entities (functions, constants) and definitions
of internal items (functions, variables, constants). All data is explicitly
typed. The name and type of external entities must be declared.  All names,
e.g. non-reserved identifiers, shall start with \llvm{@}, when they are global,
or\llvm{\%}, when they are local. For instance, \llvm{@max = external constant
  i32} declares \llvm{@max} as a 32-bit integer constant and \llvm{declare void
  @inc(i32*)} declares \llvm{@inc} as a function having as parameter a pointer
to an integer and \llvm{void} as a return type.

The type system contains the empty type \llvm{void}, an infinite, countable
number of integer types, one for each possible bit width (e.g. \llvm{i8} is the
type for 8-bit integers), and type constructors pointer (monadic \llvm{$\cdot$
  *}) and structure (polyadic \llvm{\{$\cdots$\}}). For instance \llvm{\{ i8*,
  i8, i8 \}} is the type for structures with three fields, the first having as
type pointer to \llvm{i8}. Grammar rule \nt{type\_def} states how types are
named, e.g.: \llvm{\%T1 = type \{i32, i32\}} and \llvm{\%T2 = type \{\%T1 *,
  \%T1 *\}}. Note that there is no \llvm{void *} type in the LLVM, instead
pointer values are of integer types.

Local entities are constants, variables or functions. An example of constant
definition is \llvm{@secret = constant i32 42}, it is composed of a name, type
and value. The elements of a variable definition are the name, type and code
generation attributes, e.g \llvm{@count = common global i32 zeroinitializer}.
Attributes \llvm{common}, \llvm{global} and \llvm{zeroinitializer} provide
information for target code generation: linkage type (the variable is merged
with other of the same name), the scope (global) and initialization of the
variable (zeroed is mandatory for common linkage). LLVM defines other attributes
for these aspects, but they are not needed in this work and their presentation
is omitted. For each such definition, a memory block is statically allocated to
stores the variable value. Function definitions are composed of the signature
and body. The signature contains the return type, name, parameters, and
attributes for target code generation. Attribute \llvm{nounwind} states that the
function never generate exceptions. The body is a sequence of blocks of
instructions in single-static assignment (SSA) form, i.e. are such that each
variable may be assigned exactly once.

The grammar rule \nt{inst} describes the different kind of instructions. All
instructions producing a value assign it to a fresh variable.  Instruction
\llvm{alloca} allocates a memory block the size of the given type on the stack
segment. This memory is automatically freed when the current frame is popped
from the stack. The LLVM does not include instructions for allocating from the
heap. Arithmetic operations are binary, and include signed quotient and
reminder. Arithmetic comparisons compare two values according to the first
operand and return a 1-bit integer value.  Instruction \llvm{call} invokes the
given function with the given arguments and assigns the result to a fresh
variable, which is of the given type.  In general, \llvm{getelementptr} gets the
address of an element in an aggregate object through indexing. In the target
LLVM subset, aggregate objects are structures, the first index has type
\llvm{i32} and value \llvm{0} to select the first structure at the given
location \nt{exp}, and the second index selects a field in that structure.
Instruction \llvm{load} assigns to a fresh variable \nt{name} the contents of a
memory address of type \nt{type} specified by \nt{exp} (e.g. in
fig.~\ref{fig:ex-llvm}).  Instruction \llvm{store} writes a value to memory
address (again, see fig.~\ref{fig:ex-llvm}).  Instruction \llvm{br} has two
forms.  In the first form, it directs conditionally the control flow to one of
two blocks.  If the given expression evaluates to \llvm{1}, control goes to the
first block, otherwise to the second. In the second form, the control jumps
unconditionally to the given block.
% Instruction \llvm{switch} directs the
% control flow to one of several blocks. The given expression is evaluated and
% control goes to the first branch with the matching value, or to the default
% branch (appearing first in the instruction).  
Finally, instruction \llvm{ret}
ends the current function call, optionally returning a value.

The expression language is limited to names (local and global), integer literals
and selection of an element in a structure.

To conclude, we make no assumption on the existence of a library to obtain
resources managed by the operating system, such as dynamic memory allocation. In
the absence of such resource, all data must be allocated either statically, or
on the current stack frame, by using the \llvm{alloca} instruction.

\section{Translation: a global picture}
\label{sec:overview}

In addition to producing LLVM code that is faithful to the design developed in
B, our code generator is designed to satisfy the following two requirements:
\begin{enumerate}
\item Memory is allocated statically, save for the function call frame stack: no
  dynamic memory allocation library is required. Many safety critical systems
  preclude use of this programming facility.
\item The code generator features separate compilation: an internal change in a
  component should only require generating new LLVM code for that component, and
  not for the other components that depend on it. This condition is important
  for large, distributed projects.
\end{enumerate}

In Atelier-B, software development is organized in libraries of modules. A
module has a specification, called a machine. A machine may be:
\begin{description}
\item[basic] In that case its implementation is carried out outside the realm of
  the B method. A possible application is a set of related I/O routines. Even
  though no LLVM code needs to be generated to implement the basic machines, the
  code generator needs still to produce LLVM declarations of the entities
  (e.g., constants and operations) that are supplied by those machines.
\item[developed] In that case, this module also contains a series of
  refinements, eventually reaching an implementable version, from which LLVM
  code may be generated.
\end{description}
Amongst the several constructs of the B notation supporting modular design, the
\emph{import} relation is related to separate compilation. At the implementation
level, one module may import several instances of a machine to construct its
internal data structures. Such imported machines may be either basic, or
developed. In the latter case, the implementation of the imported machines may
in turn import other machine instances and there is no pre-established limit to
such chain of imports. Hence, the import relation between module instances forms
a tree, where the root is an implementation and the descendants of a node are
the instances imported from the module in that node.

\begin{figure}
\begin{center}
\includegraphics[width=.8\textwidth]{b-project.pdf}
\caption{Structure of the imports relation in a B project}
\label{fig:structure}
\end{center}
\end{figure}
Figure~\ref{fig:structure}
illustrates this with the structure of a B project where an implementation
\B{I0} of a specification \B{M0} imports one unnamed instance of a basic machine
\B{M2} and two instances named \B{A} and \B{B} of a developed machine \B{M1}. In
turn its implementation \B{I1} itself imports two instances \B{C} and \B{D} of a
developed machine \B{M3}, implemented as \B{I3} (which has no imports itself).
The generation of LLVM code for the project shall allocate memory for one
instance of \B{I0}, two instances of \B{I2} and four instances of \B{I3}.

At that point, let us remark that, when generating code for a given module, we
need to consider the following two cases:
\begin{enumerate}
\item The code generation aims at producing a component, that will further be
  instantiated by importations. We call this the \emph{COMP} mode for code
  generation. This case is further split into two sub-cases:
  \begin{enumerate}
  \item The module is basic, and only declarations need to be emitted.
  \item The module is developed, and declarations and definitions are produced.
  \end{enumerate}
\item The generated code corresponds to a full project. Given the root module,
  the code generator must identify all the transitively imported module instances. We call this the \emph{PROJ} mode for code generation.
\end{enumerate}
One key issue in the code generation is how memory shall be allocated to store
the state of the modules. To satisfy our first initial requirement (no dynamic
memory allocation), this allocation shall be static, that is to say the
generator must include the code that creates the module instances when the
project starts execution. Note that this cannot be done in \emph{COMP} mode as,
at that point, we cannot assume any knowledge on the number of instances the
module being compiled we there be at run time. Consequently, the responsibility
for the static allocation of instances is assigned to the \emph{PROJ} mode,
which shall also perform the proper bindings between the actual and formal
instances in the generated code.

To encode the data space of a module, the translation produces a structure type,
named \llvm{\%M\$state\$}, such that the elements of the structure represent the
variables and the instances of imported modules. To satisfy our second initial
requirement (separate compilation), the instances cannot be part of the
structure itself, instead references to those instances (encoded as pointers)
are used. Let \llvm{\%M\$ref\$} denote the type pointer to \llvm{\%M\$state\$}.
To encode behavior of a module, each operation \B{op} is translated into the
definition of a function named \llvm{@M\$op}The parameter list of the functions
thus produced contains one item for each input and output of the corresponding
operation. There is also one parameter of type \llvm{\%M\$ref\$}, to pass the
address of the structure encoding instance associated with that operation. Also
the translation produces a special function \llvm{@M\$init\$} responsible for
the initialization. The parameters of this function are the addresses of the
instances found in the import tree of the module (including the module itself).

Our approach consists in dividing the translation into three sections, and to
combine these sections differently in \emph{PROJ\/} and \emph{COMP\/} mode. For
a given module \B{M}, the sections are the following:
\begin{description}
\item[typedef:] Defines an LLVM data type \llvm{\%M\$state\$} corresponding to
  the data space of the module. In case the data space of a module is empty,
  this section is empty and the type is not defined:
  \begin{tabbing}
    \llvm{\%M\$state\$ = type \{ \ListOf{\nt{type}} \}}
  \end{tabbing}
\item[interface:] Defines an LLVM type \B{M\_ref}, pointer to \B{M\_data}, if
  this has been defined. 
  \begin{tabbing}
    \llvm{\%M\$ref\$ = type \%M\$state\$*}
  \end{tabbing}
  The interface also contains declarations of LLVM
  functions: one implementing the initialisation of the module, and one for each
  operation of the module. If the data space of the module is empty, no function
  definition corresponding to the initialisation is emitted
  \begin{tabbing}
    foo \= \kill
    \llvm{declare void @M\$op(\%M\$ref\$, \ListOf{\nt{type}})} \\
    \llvm{declare void @M\$init\$(\%M\$ref\$, \ListOf{\nt{type}})}
  \end{tabbing}
\item[implementation:] Defines the functions declared in the interface. This
  section is produced only for the developed modules.
  \begin{tabbing}
    foo \= \kill
    \llvm{define void @M\$op(\%M\$state\$* \%self\$, \ListOf{\nt{param}}) \{} \\
    \> \llvm{\ListOf{\nt{block}}} \\
    \> \llvm{exit: ret void} \\
    \llvm{\}}
  \end{tabbing}
\end{description}
Assuming these three sections are available, the \emph{COMP\/} translation
assembles them as follows:
\begin{center}
  \begin{tabbing}
    foo \= foo \= \kill
    \textit{for each imported module} \B{Q} \textit{generate} \\
    \> \textit{generate a type declaration for the state space of} \B{Q} \\
    \> \llvm{\%Q\$state\$ = type opaque} \\
    \> \textit{include the interface section of }\B{Q} \\
    \textit{include the typedef section of} \B{M} \\
    \textit{include the implementation section of} \B{M}
  \end{tabbing}
\end{center}
For a module \B{M}, the \emph{PROJ\/} translation is organized as follows:
\begin{center}
  \begin{tabbing}
    foo \= foo \= \kill
    \textit{for each transitively imported component} \B{Q} \\
    \> \textit{include the typedef section of \B{Q}} \\
    \textit{for each instance} \B{Q} \textit{imported transitively through} \B{path} \\
    \> \textit{declare a variable of type} \llvm{\%Q\$state\$}: \\
    \> \llvm{@Q$\lbrack$path$\rbrack$ = common global \%M\$state\$ zeroinitializer} \\

    \textit{for each imported component} \B{Q} \\
    \> \textit{include the interface section of} \B{Q} \\
    \textit{define a function} \llvm{\%\$init\$} \textit{with a call to the } \\
    \> \textit{initialization function of} \B{M} \textit{with the proper bindings} \\
    \> \llvm{define void @\$init\$(void) \{} \\
    \> \> \llvm{call void @M\$init\$(@M, \{ \ListOf{instances} \}) \{} \\
    \> \> \llvm{exit: ret void} \\
    \> \llvm{\}} \\
    \> 
  \end{tabbing}
\end{center}

The detailed presentation of the translation of the different elements of a B
implementation is divided into:
\begin{description}
\item[Modularity] constructs to combine B components are described in
  sec.~\ref{sec:module}.
\item[Data] translation is briefly described sec.~\ref{sec:data}, including
  the encoding of B data types.
\item[Expressions] and conditions are codified as LLVM instruction blocks, as
  described in section~\ref{sec:expr}.
\item[Control] flow, instructions are translated to LLVM instructions as
  discussed in sec.~\ref{sec:control}, which also presents the encoding
  of B operations and initialization.
\end{description}

\paragraph{Notations.} Code generation is defined with a set of rules defined by
structural recursion. For each grammatical construct, a code generation function
is defined: $\trad{~}{\B{Oper}}$ for operations, $\trad{~}{\B{Inst}}$ for
instructions, $\trad{~}{\B{Expr}}$ for expressions and so forth. Each such
function usually produces a text corresponding to bits of LLVM source code. It
may have additional parameter and results. The symbol $\sep$ is used to clearly
separate composed results. Auxiliary functions are introduced for
conveniency. To abbreviate the presentation, we denote
$\trad{~}{\ListOf{\B{A}}}$ as the translation function for sequences of \B{A}
elements, defined as the concatenation of the application of $\trad{~}{\B{A}}$
to the elements of the sequence.

In those rules, $\PH{t}$ indicates a text place-holder where the value of $t$
shall be inserted and $\nl$ denotes a new line. For instance, we define
$\local(\B{n}) \isdef \llvm{"\%\PH{\B{n.id}}"}$ as a function that given an
element \B{n} of the abstract syntax tree returns the corresponding LLVM local
name, obtained by prepending character \% to its identifier; and $\Global(\B{n})
\isdef \llvm{"@\PH{\B{n.root.id}}\$\PH{\B{n.id}}"}$ yields the corresponding
LLVM global name, composed by special characters (to avoid name conflicts), the
name of the component where it is defined and its own name.  Also, we define
three auxiliary functions responsible to produce such identifiers:
\begin{align*}
\state(\B{n}) & \isdef \llvm{"\%\PH{\B{n.id}}\$state\$"} & \text{ (the type for the state)} \\
\init(\B{n}) & \isdef \llvm{"@\PH{\B{n.id}}\$init\$"} & \text{ (the initialization function)} \\
\self(\B{n}) & \isdef \llvm{"@\PH{\B{n.id}}\$self\$"} & \text{ (the instance in a top-level module)}
\end{align*}
We assume an unlimited supply of labels and local variables: we use $\lbl$ to
return a fresh LLVM label and $\variable$ for a fresh local variable name.

Finally, we need a function that, given a concrete variable, or an imported
machine, returns the position of the corresponding element in the structure
representing the state of the implementation. The position of the elements in
the structure is the same as the position of corresponding B construct in the
source code:
\begin{align*}
  \idx(\B{n}) & \isdef \mathit{index}(\B{n}, \B{n.root.impo} \conc
\B{n.root.vari})
\end{align*}
where $\mathit{index}(e, s)$ is a function that returns the position of $e$ in
sequence $s$ and $\conc$ is the concatenation operator for sequences.

\section{Translation: Modularity Aspects}
\label{sec:module}

[[[ THIS SECTION IS DEPRECATED ]]]

The translation of a B implementation to LLVM, following the template sketched
in Figure~\ref{fig:skel}, is specified by the following rule, where \B{n} is the
root of the implementation representation and $\top^?$ a boolean indicating if
\B{n} is the implementation of a top-level component or of a library machine:
\begin{align*}
  \lefteqn{\trad{\B{n}, $\top^?$}{\B{Impl}} \isdef } \\
  & \llvm{"\PH{\trad{\B{n.impo}, $\emptyset$}{\ListOf{\B{State}}}_{def}}} & \text{defines the type of imported modules} \\
  & \llvm{\state(\B{n}) = type \PH{\trad{n}{State}_{expr}}} & \text{defines the type of current module} \\
  & \llvm{\PH{\trad{\B{n.impo}, $\emptyset$}{\ListOf{\B{Bhv}}}_{decl}}} & \text{declares of imported modules operations} \\
  & \llvm{\PH{\trad{\B{n.init}}{\B{Init}}}} & \text{defines the initialization function} \\
  & \llvm{\PH{\trad{\B{n.op}}{\ListOf{\B{Oper}}}}} & \text{defines functions implementing operations}\\
  & \llvm{\PH{\trad{\B{n}, $\top^?$}{\B{Inst}}}"} & \text{creates instance (conditional)}
\end{align*}

To handle the importation of a machine, one needs to produce the definition of
the type for the state of its implementation. Since this implementation may reference
other modules, the definition is recursive (parameter $A$ is an accumulator to avoid
duplicating such definitions):
\begin{align*}
  \trad{\B{n}, $A$}{\B{State}}_{def} \isdef 
  & \IF \B{n} \not\in A \THEN \\
  & \quad \LET \B{m} = \B{n.mach} \IN \\
  & \quad \quad \LET \B{i} = \B{m.impl} \IN \\
  & \quad \quad \quad \llvm{"\PH{\trad{\B{i.impo}, $A \cup \{ \B{n} \}$}{\ListOf{\B{State}}}_{def}}} \\
  & \quad \quad \quad \llvm{\PH{\state(\B{m})} = type \PH{\trad{i}{State}_{expr}} \nl"}
\end{align*}

The following rules specify how the type representing the state of an
implementation is generated. It is a structure type, with one element per
imported machine, and one element per variable. The definition of such type is
just the enumeration of the type of its elements (the individual elements are
referenced, not by name, but by position inside the structure):
\begin{align*}
  \trad{\B{n}}{\B{State}}_{expr} \isdef & \llvm{"\{ \PH{\trad{\B{n.impo}}{\ListOf{\B{Impo}}}_{expr}},  \PH{\trad{\B{n.var}}{\ListOf{\B{Vari}}}_{expr}} \}"} \\
  \trad{\B{n}}{\B{Impo}}_{expr} \isdef & \state(\B{n.mach}) \\
  \trad{\B{n}}{\B{Vari}}_{expr} \isdef & \trad{\B{n.type}}{\B{Type}}
\end{align*}

The code generation for the functions implementing the initialization and
operations of imported machine is specified by the following rules. Not that
this definition needs not be recursive, as only the operations of a directly
imported machine may be called in an implementation. Again, parameter $A$ plays
the role of an accumulator to avoid (harmless, but useless) duplicate
declarations.
\begin{align*}
  \trad{\B{n}, $A$}{\B{Bhv}}_{decl} \isdef 
  & \IF \B{n} \not\in A \THEN \\
  & \quad \LET \B{m} = \B{n.mach} \IN \\
  & \quad \quad \LET \B{i} = \B{m.impl} \IN \\
  & \quad \quad \quad \llvm{declare void \PH{\init(\B{m})}(\PH{\state(\B{m})} *) \nl} \\
  & \quad \quad \quad \llvm{\PH{\trad{\B{m.op}}{\ListOf{\B{Oper}}}_{decl}}"} \\
  \trad{\B{n}}{\B{Oper}}_{decl} \isdef &
  \llvm{"declare void \PH{\Global(\B{n})}(\PH{\state(\B{n.root})}, \PH{\trad{\B{n.inp}}{\ListOf{\B{Inp}}}}, \PH{\trad{\B{n.out}}{\ListOf{\B{Out}}}})"}\\
  \trad{\B{n}}{\B{inp}}_{decl} \isdef & \trad{\B{n.type}}{\B{Type}} \\
  \trad{\B{n}}{\B{out}}_{decl} \isdef & \llvm{"\PH{\trad{\B{n.type}}{\B{Type}}}*"}
\end{align*}

The last rule is responsible for producing the code creating the instance when the
implementation is that of the top-level component of a design:
\begin{align*}
\lefteqn{  \trad{\B{n}, $\top^?$}{\B{Inst}} \isdef } \\
  & \IF \top^? \THEN \llvm{"\PH{\self(\B{n})} = common global \PH{\state(\B{n})} zeroinitializer \nl"} \\
  & \ELSE \llvm{""}
\end{align*}

\section{Translation: Data}
\label{sec:data}

The translation presented in this paper is currently limited to integers and
booleans. In the case of integers, the translation is based on the supposition
that the B0 definition of \B{INT} is the range $-2^{31}.. 2^{31}-1$. We have
$\trad{BOOL}{\B{Type}} \isdef \llvm{i1}$ and $\trad{INT}{\B{Type}} \isdef
\llvm{i32}$.

B constants shall be translated to constants in the LLVM program. The
translation of scalar constants consists in substituting them with their value.
\DD{The translation of non-scalar constants will be defined in a next version of
  this document.}  The definition of the translation for expressions is
described next in section~\ref{sec:expr}.

\section{Translation: Expressions}
\label{sec:expr}

To evaluate a composed expression using the LLVM, one must first generate a
sequence of instructions to evaluate each argument, the resulting value being
stored in a temporary variable, and second generate an instance of the LLVM
instruction corresponding to the expression argument.  Translation of
expressions is defined as functions that take as input some class of expressions
and produce a triple of strings $p \sep v \sep t$, where $p$ is a possibly empty
sequence of instructions required in the evaluation (e.g.  for the
sub-expressions), $v$ is the LLVM expression representing the resulting value,
and $t$ represents the LLVM type of the expression. In the case of composed
predicates, though, the evaluation is different as boolean operations need to be
implemented with conditional and unconditional jumps.

To facilitate the defintiion of rules that apply to similar cases, we introduce
mapping $\opmap$, that associates B operators to corresponding LLVM keywords:
$$\opmap = \{ 
\begin{array}[t]{l}
  = \mapsto \llvm{"eq"}, 
  \ne \mapsto \llvm{"ne"}, 
  < \mapsto \llvm{"slt"}, 
  \le \mapsto \llvm{"sle"}, 
  > \mapsto \llvm{"sgt"}, 
  \ge \mapsto \llvm{"sge"}, \\
  + \mapsto \llvm{"add"} \}.
  \end{array}
$$

\paragraph{Atomic expressions} have a simple translation. In the case of
literals, one only needs to issue the corresponding LLVM literal:
$\trad{TRUE}{\B{Lit}} \isdef \llvm{""} \sep \llvm{"1"} \sep \llvm{"i1"}$,
$\trad{FALSE}{\B{Lit}} \isdef \llvm{""} \sep \llvm{"0"} \sep \llvm{"i1"}$ and
$\trad{l}{\B{Lit}} \isdef \llvm{""} \sep \llvm{"\PH{l}"} \sep \llvm{"i32"}$ when
$\B{l.type} = \B{INT}$.

For identifiers, three cases are possible: global constants, local variables,
and module state variables. For state variables, the translation issues
statements to assign the address of the corresponding element in the state
structure to pointer variable $p$ and then dereference $p$ to assign $v$:
\begin{align*}
\lefteqn{\trad{n}{\B{Name}} \isdef \LET t = \trad{n.type}{\B{Type}} \IN} \\
& \IF \B{n} \mbox{ is a constant } \THEN \trad{n.value}{\B{Expr}} \\
& \ELSIF \B{n} \mbox{ is a local variable } \THEN 
\llvm{""} \sep \local(\B{n}) \sep t \\
& \ELSE (\B{n} \mbox{ is a state variable }) \LET p = \variable \AND v = \variable \IN \\
& \quad \llvm{"\PH{p} = getelementptr \PH{\state(\B{n.root})} \%self\$, i32 0, i32 \PH{\idx(\B{n})} \nl} \\
& \quad \llvm{\PH{v} = load \PH{t}* \PH{p}\nl"} \sep v \sep t
\end{align*}

\paragraph{Comparisons} are formed by the application of a relational operator
to two expressions.

The code responsible for evaluating a comparison \B{r} is defined by the
following rule. First both arguments are evaluated in $p_1$ and $p_2$, and then
an instance of the LLVM comparison instruction \llvm{icmp} is used to compare
the temporary variables $v_1$ and $v_2$, of type $t_1 = t_2$, storing the result
in the fresh variable $v$.
\begin{align*}
\lefteqn{\trad{r}{\B{Comp}} \isdef
  \LET
  p_1 \sep v_1 \sep t_1 = \trad{r.arg1}{\B{Expr}} \AND
  p_2 \sep v_2 \sep t_2 = \trad{r.arg2}{\B{Expr}}} \\
& \IN \LET v = \variable \IN \\
& \llvm{"\PH{p_1}\quad\PH{p_2}\quad\PH{v} = icmp \PH{\opmap(\B{r.op})} \PH{t_1} \PH{v_1} \PH{v_2} \nl"} \sep v
\end{align*}

\paragraph{Arithmetic operations}\DD{Actually some arithmetic operations are not binary and different translation rules shall be written.} have a similar 
structure and interpretation as comparisons. Here is the rule for binary
arithmetic operations:
\begin{align*}
\lefteqn{\trad{r}{\B{Term}} \isdef
  \LET
  p_1 \sep v_1 \sep t_1 = \trad{r.arg1}{\B{Expr}} \AND
  p_2 \sep v_2 \sep t_2 = \trad{r.arg2}{\B{Expr}}} \\
& \IN \LET v = \variable \IN \\
& \llvm{"\PH{p_1}\quad\PH{p_2}\quad\PH{v} = i32 \PH{\opmap(\B{r.op})} \PH{t_1} \PH{v_1} \PH{v_2} \nl"} \sep v \sep \llvm{i32}
\end{align*}

Next is the translation of the unary \B{succ} operator:
\begin{align*}
  \trad{n}{\B{succ}} \isdef & \textbf{ let } p \sep v \sep t = \trad{n.arg}{\B{expr}} \IN  \\
  & \quad \LET w = \variable \IN \\
  & \quad \quad \llvm{"\PH{p}} \\
  & \quad \quad \llvm{\PH{w} = add i32 1, \PH{v} \nl"} \sep w \sep \llvm{i32}
\end{align*}

\paragraph{Predicates}
The translation of conditions takes as input a condition node \B{n}, as well as
two labels $\ell_t$ and $\ell_f$ that correspond to program locations where the
execution shall jump when the condition is evaluted to true or false,
respectively.  The first rule defines the translation of an atomic relation:
\begin{align*}
  \lefteqn{\trad{n, $\ell_t$, $\ell_f$}{\B{Form}} \isdef
  \textbf{ let } p \sep v = \trad{n}{\B{Comp}} \IN} \\
  & \llvm{"\PH{p} \quad br i1 \PH{v}, label \%\PH{\ell_t}, label \%\PH{\ell_f} \nl"}
\end{align*}

\noindent The translation for negation is simply: $\trad{not f, $\ell_t$,
  $\ell_f$}{\B{not}} \isdef \trad{f, $\ell_f$, $\ell_t$}{\B{Form}}$.  The
following rule handles the case of conjunctions. Note that it produces code with
a ``short-cut'' when the evaluation of the first argument yields false:
\begin{align*}
\begin{split}
  \lefteqn{\trad{n, $\ell_t$, $\ell_f$}{\B{and}} \isdef \LET \ell = \lbl \IN} \\
  & \LET p_1 = \trad{n.arg1, $\ell$, $\ell_f$}{\B{Form}}
  \AND p_2 = \trad{n.arg2, $\ell_t$, $\ell_f$}{\B{Form}} \IN \\
  & \quad \llvm{"\PH{p_1} \PH{\ell} :  \nl \PH{p_2}"}
\end{split}
\end{align*}
The definition of the translation for disjunctions is similar and is omitted.

\section{Translation: Control Flow and Instructions
\label{sec:control}}

This section addresses how operations in B implementations are translated to
LLVM functions. It also discusses initialization, which might be seen as a
special kind of operation. For operations, the presentation of the translation
is divided into three steps: the signature, the local data, and the body
instructions.

The signature of the function is composed of the name of the function, the
result type, the name and type of the parameters. The signature may carry other
information such as linkage, visibility, calling convention, etc. The B parser
provides this information after the type checking phase has been completed.  On
the one hand LLVM functions may only have one result, on the other hand B0
operations may produce several results. The translation of output parameters
will be function parameters passed by reference, i.e. they are pointers. For
uniformity of treatment, the return type of the produced LLVM functions is
\llvm{void}. Translation of operation headers is defined in
sec.~\ref{sec:trad-header}. The memory to store local variables is allocated on
the stack, using the \llvm{alloca} instruction. We describe a function in
sec.~\ref{sec:trad-alloc} that, applied to a B operation, returns the name and
type of its local variables used in the operation. These names are translated to
conform to LLVM restrictions. The body of the operation has to be transformed to
a sequence of LLVM blocks of statements. For each kind of instruction one (or
more) translation function is defined. The details are provided in
section~\ref{sec:trad-instr}.  The results of the translation of each aspect are
combined as follows:
\begin{align*}
\begin{split}
  \lefteqn{\trad{\B{op}}{\B{oper}} \isdef} \\
  & \LET h = \trad{\B{op}}{sig} \AND   a = \trad{\B{op.body}}{alloc} \AND i = \trad{\B{op.body}, \llvm{exit}}{\ListOf{\B{Inst}}}_L \IN  \\
  & \llvm{"define void \PH{h} \{\nl} \\
  & \quad \llvm{entry:\nl} \\
  & \quad \quad \PH{a} \\
  & \quad \quad \PH{i} \\
  & \quad \llvm{exit: ret void\nl} \\
  & \llvm{\}\nl"}
\end{split}
\end{align*}

Initialization of a B implementation shares many characteristics with operations
but needs to deal with initialization of imported modules. It is discussed in
sec.~\ref{sec:trad-init}.

\subsection{Sub-routines signatures}
\label{sec:trad-header}

The following rule specifies the translation of operation headers. Its
parameters are the operation identifier, inputs, outputs, the name of the
machine it belongs to, and the results are the LLVM function definition
header: \\
\noindent$\trad{\B{o}}{sig} \isdef \llvm{"\PH{\Global(\B{o})}(\PH{\state(\B{o.root})}* \%self\$, \PH{\trad{\B{o.inp}}{\ListOf{inp}}}, \PH{\trad{\B{o.out}}{\ListOf{out}}})"}$

Next are the rules defining the translation of individual input and output
parameters. They have local scope and they are translated by prepending
\llvm{\%}:
\begin{align*}
  \trad{i}{inp} \isdef 
  \llvm{"\PH{\trad{i.type}{\B{Type}}} \PH{\local(\B{i})}"} \quad \quad
  \trad{o}{out} \isdef 
  \llvm{"\PH{\trad{o.type}{\B{Type}}}* \PH{\local(\B{o})}"}
\end{align*}

\subsection{Stack allocation}
\label{sec:trad-alloc}

B operations may have local variables, in which case the translator needs to
issue LLVM stack allocation instructions. Such instructions have two arguments:
the type of values to store, and a temporary variable that is assigned the
address (see figure~\ref{fig:llvm-grammar} for the concrete syntax). The
generation of stack allocation code visits the parse tree of the operation body
in a pre-order depth-first traversal, generating one statement for each local
variable declaration found. The detailed definition of this traversal is omitted
for space reasons. For each B0 variable \B{v} found, a LLVM name is thus
created.
\begin{align*}
  \trad{v}{alloc} \isdef 
  \llvm{"\PH{\local(\B{v})} = alloca \PH{\trad{v.type}{\B{Type}}} \nl"}
\end{align*}

After the stack allocations, the LLVM source code contains the statements that
encodes the instructions in the body of the operation, as described in the
following section.

\subsection{Translation of instructions}
\label{sec:trad-instr}

One issue in the translation of instructions is the disruption of the flow graph
intro a linear sequence of instructions and branches. This requires the creation
of instruction labels. As a consequence, our definition of the translation of
instructions has two kinds of rules, those that take a label as parameter, and
those that do not. Such label parameter indicates where the control should go
after the corresponding execution is executed.

\begin{align*}
\trad{\B{i}, $\ell$}{\B{Inst}}_L \isdef
& \IF \B{i} \mbox{ is \B{If}} \THEN \trad{\B{i}, $\ell$}{\B{If}}_L \\
& \ELSIF \B{i} \mbox{ is \B{While}} \THEN \trad{\B{i}, $\ell$}{\B{While}}_L \\
& \ELSIF \B{i} \mbox{ is \B{Blk}} \THEN \trad{\B{i.body}, $\ell$}{\ListOf{\B{Inst}}}_L \\
& \ELSE \LET p = \trad{\B{i}}{\B{Inst}} \IN \\
& \quad \llvm{"\PH{p}} \\
& \quad \llvm{ br label \%\PH{\ell} \nl"}
\end{align*}

Function $\trad{\B{il}, $\ell$}{\ListOf{\B{Inst}}}_L$ takes as input a sequence
of B instructions \B{il} and a label $\ell$ and produces the source code of a
LLVM block with $\ell$ as exit point. It is defined as follows:

\begin{align*}
\trad{\B{il}, $\ell$}{\ListOf{\B{Inst}}}_L \isdef & \IF \B{il} \mbox{ is empty} \THEN \llvm{"br label \%\PH{\ell} \nl"} \\
& \ELSE \\
& \LET \B{i}, \B{il'} = \B{il}  \IN \\
& \quad \quad \LET p_1 = 
\begin{array}[t]{l}
  \IF \B{i} \mbox{ is \B{If} or \B{While}} \THEN \\
  \quad \LET \ell' = \lbl \IN \llvm{"\PH{\trad{i, $\ell'$}{\B{Inst}}_L} \PH{\ell'}: \PH{p}"} \\
  \ELSE \llvm{"\PH{\trad{i}{\B{Inst}}} \PH{p}"} \IN
\end{array} \\
& \quad \quad \IN \\
& \quad \quad \quad \llvm{"\PH{p_1}} \\
& \quad \quad \quad \llvm{\PH{\trad{il', $\ell$}{\ListOf{\B{Inst}}}_L}"}
\end{align*}

Eventually, $\trad{}{\B{Inst}}$ defines the translation for individual
instructions. Its definition is specialized for each class of instructions as
detailed in the rest of this section. Note that a new block is created after
each \B{If} and \B{While} instruction.

\paragraph{Becomes equal instructions} must evaluate the target and the source
of the assignment, and then copy the result of the latter in the former, using
the LLVM instruction \llvm{store}. The evaluation of the target is defined with
$\trad{}{lv}$, which yields the corresponding LLVM source code $l$ and the
variable $p$ containing the assigned location. The evaluation of the source
yields the corresponding code $r$, the variable $v$ holding the result, and its
type $t$.
\begin{align*}
\begin{split}
  \trad{a}{\B{Beq}} \isdef 
  & \LET l \sep p \sep t'= \trad{i.lhs}{lv} \AND r \sep v \sep t = \trad{i.rhs}{\B{Expr}} \IN \\
  & \llvm{"\PH{l}} \\
  & \llvm{\PH{r}} \\
  & \llvm{store \PH{t} \PH{v}, \PH{t'} \PH{p} \nl"}
%   & l \conc r \conc 
%   \llvm{"store "} \conc t \conc \texttt{" "} \conc v \conc \llvm{", "}
%   \conc t \conc \llvm{"* "} \conc p \conc \nl
\end{split}
\end{align*}
The assignment target may be either local to the operation or a state
variable. In the former case, it can be referenced by the corresponding LLVM
identifier. In the latter case, it is represented as an element of the state
structure $\Sigma$, and its address must be calculated and assigned to a new temporary
variable.
\begin{align*}
\lefteqn{\trad{n}{lv} \isdef} \\
& \LET t = \llvm{"\PH{\trad{\B{n.type}}{Type}} *"} \IN \\
& \quad \IF \mbox{\B{n} is a local variable} \THEN \llvm{""} \sep \local(\B{n}) \sep t \\
& \quad \ELSE \LET v = \variable \IN \\
& \quad \quad \llvm{"\PH{v} = getelementptr \PH{\state(\B{n.root})}* \%self\$,} \\
& \quad \quad \quad \llvm{i32 0, i32 \PH{\idx(\B{n})} \nl"} \sep v \sep t
\end{align*}

\paragraph{Call up} instructions apply B operations with some values and obtain
results. The LLVM code generator produces instructions to evaluate the operation
arguments, the address of the variables receiving the results as well as an
instance of the LLVM \llvm{call} instruction. In addition, the proposed
translation includes as argument the address of a structure storing the state of
the component corresponding to this operation. Note that this component may be
either the implementation being translated (in the case of a so-called local
operation), or an instance of an imported machine.
\begin{align*}
  \trad{c}{\B{Call}} \isdef 
  & \LET p_i \sep il = \trad{c.inp}{\ListOf{Inp}} \AND p_o \sep ol = 
  \trad{c.out}{\ListOf{Out}} \AND \\
  & \quad id = \Global(\B{c}) \AND ty = \llvm{"\PH{\state(\B{c.root})} *} \AND \\
  & \quad val \sep p = \IF \B{c.inst} = \text{None} \\
  & \quad \quad \THEN \llvm{"\%self\$"} \sep \llvm{""} \\
  & \quad \quad \ELSE \LET v = \variable \IN \\
  & \quad \quad \quad v \sep \\
  & \quad \quad \quad \llvm{"\PH{v} = getelementptr \PH{ \state({\B{c.inst.root}})} \%self\$, i32 0, i32 \PH{\idx(\B{c.inst})} \nl"} \\
  & \IN \\
  & \quad \llvm{"\PH{p_i}} \\
  & \quad \llvm{\PH{p_o}} \\
  & \quad \llvm{\PH{p}} \\
  & \quad \llvm{call void \PH{id}(\PH{ty}* \PH{val}, \PH{il}, \PH{ol}) \nl"}
\end{align*}
The translation of input and output arguments produces two code snippets. The
first corresponds to the LLVM instruction sequence $p$ responsible for
evaluating the argument, and the second to the type $t$ and name $v$ of the
variable storing the resulting value:
\begin{align*}
  \trad{i}{Inp} \isdef 
  & \LET p \sep v \sep t = \trad{i}{\B{Expr}} \IN p \sep \llvm{"\PH{t} \PH{v}"}\\
  \trad{o}{Out} \isdef 
  & \LET p \sep v \sep t = \trad{o}{lv} \IN p \sep \llvm{"\PH{t} \PH{v}"}
\end{align*}

\paragraph{Identity} does nothing and no code is generated:
\begin{align*}
\begin{split}
  \trad{n}{\B{Skip}} \isdef & \llvm{""}
\end{split}
\end{align*}

\paragraph{Block instructions} are structuring constructs that have no
operational intent; their translation is that of their body: $\trad{i,
  $\ell$}{\B{Blk}} \isdef \trad{i.body, $\ell$}{\ListOf{\B{Inst}}}$.

\paragraph{If instructions} are sequences of possibly conditional blocks of
instructions, i.e. branches. After a branch is executed, control must go
to the next block $\ell$.
\begin{align*}
  \trad{i, $\ell$}{\B{If}}_L \isdef \trad{i.branches, $\ell$}{\ListOf{\B{IfBr}}}
\end{align*}
To translate conditional branches, one must first generate code to evaluate the
condition of the branch, followed by an LLVM conditional branching statement,
translate the body of the branch, and add an unconditional branch to the block
following the corresponding \B{If} statement.  The translation for branches has
as arguments a branch \B{b}, the remaining branches \B{bl}, the lexicon, and the
label $\ell$ of the block following the \B{If} statement they belong to. The
following definition applies for the case of conditional branches:
\begin{align*}
\begin{split}
  \trad{\B{b bl}, $\ell$}{\ListOf{\B{IfBr}}} \isdef & \LET \ell_1 = \lbl \AND \ell_2 = \lbl \IN \\
  & \llvm{"\PH{\trad{b.cond, $\ell_1$, $\ell_2$}{\B{Pred}}} } \\
  & \llvm{\PH{\ell_1} : \PH{\trad{b.body, $\ell$}{\ListOf{\B{Inst}}}_L}} \\
  & \llvm{\PH{\ell_2} : \PH{\trad{bl, $\ell$}{\ListOf{\B{IfBr}}}}"} \\
\end{split}
\end{align*}
The following rule handles the case of the last branch, which might have a condition or not:
\begin{align*}
\begin{split}
  \trad{\B{b}, $\ell$}{\ListOf{\B{IfBr}}} \isdef & \IF \B{b.cond} = \text{None} \THEN
  \trad{b.body, $\ell$}{\ListOf{Inst}}_L\\
  & \ELSE \LET \ell_1 = \lbl \IN \\
  & \llvm{"\PH{\trad{b.cond, $\ell_1$, $\ell$}{\B{Pred}}} } \\
  & \llvm{\PH{\ell_1} : \PH{\trad{b.body, $\ell$}{\ListOf{\B{Inst}}}_L}}
\end{split}
\end{align*}

\paragraph{Case instructions} are formed by an expression, and a list of case
branches, each of which has a list of values and an instruction block.
Optionally, there might be a default branch, and we consider that, in this case,
the list of values is empty.

The translation to the LLVM IR uses its switch statement. It is composed of a
variable, a default jump and a jump table. The jump table is a sequence of pairs
formed by a value, and an instruction label. In the following, \llvm{\%val} is
the variable, \llvm{\%otherwise} is the default label, and the jump table has
three entries. The first corresponds to the case where the value of \llvm{\%val}
is \llvm{0}, in which case the execution jumps to the label \llvm{onzero}.
\begin{verbatim}
switch i32 %val, label %otherwise [ i32 0, label %onzero
                                    i32 1, label %onone
                                    i32 2, label %ontwo ]
\end{verbatim}

The translation of the case statement first consists in producing the code $p$
to evaluate the expression tested in the statement. When that code completes
execution, the value is stored in variable $v$ of type $t$. Then the jump table
$j$ and the code $b$ of the different branches are generated. This generation
takes also as parameters labels $\ell$ and $\ell_o$.
\begin{align*}
  \trad{v, $\ell$}{\B{Case}} \isdef & \LET p \sep v \sep t = \trad{v.expr}{\B{Expr}} \IN \\
  & \LET \ell_o = \lbl \IN \\
  & \LET j, b = \trad{v.branches, $\ell_o$, $\ell$}{\ListOf{CaseBr}} \IN \\
  & \quad \llvm{"\PH{p}} \\
  & \quad \llvm{switch \PH{t} \PH{v}, label \%\PH{\ell_o} [ \PH{j} ] \PH{b}"}
\end{align*}

The following rule defines the composition of the translation of the different
branches of a case statement. Again, this translation has two parts: bits of the
jump table $j$, and instruction blocks $b$. If a case branch has several values,
then several entries need to be created in the jump table. This is the role
of the translation function $\trad{}{\B{ListOf{CaseVal}}}$.
\begin{align*}
  \trad{b bl, $\ell_o$, $\ell_e$}{\ListOf{\B{CaseBr}}} \isdef 
  & \LET \ell = \lbl \IN \\
  & \LET j, b = \trad{bl, $\ell_o$, $\ell_e$}{\ListOf{\B{CaseBr}}} \IN \\
  & \quad \llvm{"\PH{\trad{b.values, $\ell$}{\ListOf{\B{CaseVal}}}} \PH{j}"}
  \sep \\
  & \quad
  \llvm{"\PH{\ell}: \nl \PH{\trad{b.body, $\ell_e$}{\ListOf{Inst}}_L} \PH{b}"}
\end{align*}
This rule handles the last branch of a case statement, and needs to test if
that branch is a default branch or not. If not, then one needs to be artificially
created (as this is mandatory in the LLVM-IR switch statement):
\begin{align*}
  \trad{b, $\ell_o$, $\ell_e$}{\ListOf{\B{CaseBr}}} \isdef 
  & \IF \B{b.values} = \text{None} \THEN \\
  & \quad \llvm{""} \sep \llvm{"\PH{\ell_o}: \nl \PH{\trad{b.body, $\ell_e$}{\ListOf{Inst}}_L}"} \\
  & \ELSE \LET \ell_b = \lbl \IN \\
  & \quad \trad{b.values, $\ell_b$}{\ListOf{\B{CaseVal}}} \sep \\
  & \quad \llvm{"\PH{\ell_b}: \nl \PH{\trad{b.body, $\ell_e$}{\ListOf{Inst}}_L}} \\
  & \quad \quad \llvm{\PH{\ell_o}: \nl br label \% \PH{\ell_e}}
\end{align*}

The following rules specifies that the values of a case branch are translated in
order, their translations being concatenated to form the result. The parameter
$\ell$ is the label of the instruction block containing the translation of the
branch body:
\begin{align*}
  \trad{v, $\ell$}{\ListOf{\B{CaseVal}}} \isdef & \trad{v, $\ell$}{\B{CaseVal}} \\
  \trad{v vl, $\ell$}{\ListOf{\B{CaseVal}}} \isdef & 
  \llvm{"\PH{\trad{v, $\ell$}{\B{CaseVal}}} \PH{\trad{vl, $\ell$}{\ListOf{\B{CaseVal}}}}"}
\end{align*}

This rule defines how a single value of a case branch is translated, namely
as an entry in the jump table of a switch statement:
\begin{align*}
  \trad{v, $\ell$}{\B{CaseVal}} \isdef & \LET p \sep v \sep t = \trad{v}{\B{Expr}} \IN \\
  & \quad \llvm{"\PH{t} \PH{v}, label \%\PH{\ell} \nl"}
\end{align*}

\paragraph{While instructions} are translated in two blocks of instructions:
block $\ell_1$ evaluates the loop condition and block $\ell_2$ executes the loop
body:
\begin{align*}
\begin{split}
\trad{w, $\ell$}{\B{While}} \isdef 
  & \LET \ell_1 = \lbl \AND \ell_2 = \lbl \AND c \sep v = \trad{w.cond}{\B{Pred}} \IN \\
  & \llvm{"br label \%\PH{\ell_1} \nl} \\
  & \llvm{\PH{\ell_1} : \PH{c}} \\
  & \llvm{br i1 \PH{v}, label \%\PH{\ell_2}, label \%\PH{\ell} \nl} \\
  & \llvm{\PH{\ell_2} : \PH{\trad{i.body, $\ell_1$}{\ListOf{\B{Inst}}}_L}}
\end{split}
\end{align*}

\subsection{Initialization}
\label{sec:trad-init}

[[[ DEPRECATED ]]]

Initialization is done is two steps: first all imported modules are
initialized,  second the initialization clause is executed.
\begin{align*}
  \trad{i}{Init} \isdef
  & \llvm{"define void @\PH{\B{i.id}}\$init\$() \{ \nl"} \\
  & \quad\llvm{\PH{\trad{i, i.impl}{\ListOf{InitImp}}}} \\
  & \quad\llvm{\PH{\trad{i.init, "exit"}{\ListOf{\B{Inst}}}_L}} \\
  & \quad\llvm{exit: ret void \nl} \\
  & \llvm{\}"}
\end{align*}
The initialization of an imported module is performed by a call to the
corresponding initialization function:
\begin{align*}
  \lefteqn{\trad{\B{impl}, \B{impo}}{InitImp} \isdef} \\
  & \LET v = \variable \AND \B{m} = \B{impo.mach} \IN \\
  & \quad \llvm{"\PH{v} = getelementptr \PH{\state(\B{impl.root})}* \%self\$, i32 0, i32 \PH{\idx(\B{impo})}\nl}\\
  & \quad \llvm{call void @\PH{\B{m.id}}\$init\$(\PH{\state(\B{m.impl})}*
    \PH{\ell}) \nl"}
\end{align*}

\section{Validation}
\label{sec:validation}

Before implementing this translation, we want validate the rules that define it.
For this, we follow a pragmatic approach, based on testing and simulation of the
generated LLVM artifacts, and using the existing LLVM tool set.

First, we define a sample set of small B implementation modules, that contain
all the B constructs handled in the translation. For each such module, the rules
are applied manually, yielding a LLVM module. This LLVM module is then amended
with extra code that have two purposes:
\begin{itemize}
\item a standalone executable may be generated from it;
\item the executable creates an instance of the B implementation, applies the
  initialization code, and offers to the user the choice to interactively invoke
  all the operations of the implementation; after initialization and each
  operation application the state of the implementation is printed. 
\end{itemize}

The LLVM tools that are employed are:
\begin{description}
\item[llvm-as] The LLVM assembler generates LLVM bitcode, invoked, e.g., as
  follows:
\begin{verbatim}
llvm-as-mp-3.1 example-01.llvm -o example-01.bc
\end{verbatim}
\item[lli] The LLVM bitcode interpreter, which may be invoked as follows:
\begin{verbatim}
lli-mp-3.1 example-01.bc
\end{verbatim}
\end{description}

Once a rule has been manually validated, we implement it in a prototype tool
written in Python. Unfortunately the tool does not take as input the B0 source
file, but instead a collection of Python definitions corresponding to the
syntactic entities present in the parse tree. For instance, for an integer
concrete variable named \B{value}, we would have the following Python entity:
\begin{verbatim}
value = { "kind": "Vari",
          "id": "value",
          "type": "INT",
          "scope": "impl",
          "root": counter_i}
\end{verbatim}
where \verb'counter_i' would be the Python entity representing the parse tree
root for the B implementation. Again the output generated by the prototype
is compiled and simulated using off-the-shelf LLVM tools \verb'llvm-as' and
\verb'lli'.

\section{Conclusion and future work}
\label{sec:conclusion}

This paper presents the formal specification of a translation from a large
subset of the B implementation language to LLVM, a modern compiler internal
language.  Even though this is still a work in progress, the definition is
self-contained and has a large enough scope to be applied to B implementations
where the data belongs to basic types. This specification is a blueprint for a
code synthesis tool for B that is currently being implemented. This tool will be
distributed as an extension to Atelier-B under an open-source license.

The next step is to extend the scope to the full B implementation language. This
entails that the translation must be defined for aggregate data types and
adapted to handle some syntactic sugar. We are also planning for producing a
LLVM output with debugging information. Such output would be indeed very helpful
to provide feedback to the user when applying testing to validate the produced
code.

To prove the verification of the correctness of the translation, we would have
to define the semantics of both LLVM and B in a unified framework. One possible
starting point is Vellvm~\cite{vellvm}, a framework to reason about the
correctness of LLVM programs and transformations. We would then have to extend
the framework to include B implementation language. Another possible approach
would be to translate verification conditions from the B development artifacts
as assertinons in the generated LLVM code. The compiled program would include
checks that such assertions hold while executing.

\bibliographystyle{plain}
\bibliography{b2llvm}

\newpage
\appendix

\section{Test summary}

For each test file, indicates the B0 constructs found in the implementation.
\begin{center}
\newcommand{\no}[0]{--}
\newcommand{\yes}[0]{$\circ$}
\begin{tabular}{|l|c|c|c|c|c|c|c|c|}
\hline
Example nº & 01 & 02 & 03 & 04 & 05 & 06 & 07 & 08 \\
\hline
\hline
\multicolumn{9}{|c|}{clauses} \\
\hline
import w.o prefix & \no & \no & \no & \no & \no & \no & \yes & \no \\
import w prefix & \no & \no & \no & \no & \no & \no & \no & \yes \\
variables & \yes & \yes & \yes & \yes & \yes & \yes & \no & \yes \\
constants & \no & \yes & \no & \no & \yes & \no & \yes & \yes \\
op. input & \yes & \no & \yes & \no & \no & \no & \no & \no \\
op. output & \yes & \no & \yes & \no & \no & \no & \yes & \yes \\
\hline
\multicolumn{9}{|c|}{instructions} \\
\hline
block & \yes & \no & \yes & \yes & \no & \yes & \no & \yes \\
local var & \no & \no & \yes & \yes & \no & \yes & \yes & \yes \\
identity & \no & \no & \yes & \no & \no & \no & \no & \no \\
becomes equal & \yes & \yes & \yes & \yes & \yes & \yes & \yes & \yes \\
operation call & \no & \no & \no & \no & \no & \no & \yes & \yes \\
condition & \yes & \no & \no & \no & \yes & \no & \yes & \yes \\
case & \no & \no & \no & \no & \no & \yes & \no & \no \\
assertion & \no & \no & \no & \no & \no & \no & \no & \no \\
sequence & \yes & \yes & \yes & \yes & \no & \yes & \yes & \yes \\
while & \no & \no & \no & \yes & \no & \no & \yes & \no \\
\hline
\multicolumn{9}{|c|}{types} \\
\hline
INT & \yes & \yes & \yes & \yes & \yes & \yes & \yes & \yes \\
BOOL & \yes & \yes & \no & \no & \yes & \no & \yes & \yes \\
\hline
\multicolumn{9}{|c|}{expressions} \\
\hline
int lit & \yes & \yes & \yes & \yes & \yes & \yes & \yes & \yes \\
bool lit & \yes & \yes & \no & \yes & \yes & \no & \yes & \yes \\
vars & \yes & \yes & \yes & \yes & \no & \yes & \no & \yes \\
loc. vars & \no & \no & \yes & \yes & \no & \yes & \yes & \yes \\
const & \no & \yes & \no & \no & \yes & \no & \yes & \yes\\ 
succ & \no & \yes & \no & \yes & \yes & \no & \no & \no \\
pred & \no & \yes & \no & \yes & \yes & \no & \no & \no \\
+ & \yes & \no & \no & \yes & \no & \yes & \yes & \no \\
- & \yes & \no & \no & \no & \no & \no & \yes & \no \\
* & \no & \no & \no & \no & \no & \yes & \no & \no \\
\hline
\multicolumn{9}{|c|}{formulas} \\
\hline
and & \no & \no & \no & \no & \yes & \no & \no & \no \\
or & \no & \no & \no & \no & \yes & \no & \no & \no \\
not & \no & \no & \no & \no & \yes & \no & \no & \no \\
$<$ & \yes & \yes & \no & \no & \no & \no & \yes & \yes \\
$>$ & \no & \no & \no & \no & \yes & \no & \yes & \no \\
\hline
\end{tabular}
\end{center}

\end{document}