### 2014-07-07 DOCUMENTATION/DerivedTypes.tex
Small updates about array

\subsection{Translation of structures and arrays}
\label{sec:StructArray}

Two approaches are possible to support arrays and structures. One approach is 
importing reusable libraries of AtelierB that are reported in the official 
document of ClearSy\footnote{\url{http://www.atelierb.eu/ressources/DOC/english/reusable-components-reference.pdf}}. 
This approach is supported, but the generated code is more complex and it needs 
more intermediate steps to manipulate the data of imported components. 
Moreover, this approach allows dynamic allocation for arrays, but it is 
strongly not recommended for safety-critical systems. A second approach enjoys 
the direct correspondence between B0 native types and LLVM and this approach 
is presented here. This section describes the translation of derived data 
types showed in Section~\ref{sec:data}.

\subsubsection{Structure} 
The structure declaration defines a list of variables to be placed under one 
name, allowing several variables to be accessed by one base name. This example  
initializes a structure, and gets and sets one value of xx'balance.

\begin{description}
\item[Case 1.1] Initializing a record with a sequence of expressions:
\[
  \textcolor{gray}{accountRecord :=}  \underbrace{rec(\underbrace{name:xx’name, balance:(xx’balance - amt)}_{seq\ Exp})}_{Record}
\]

\item[Case 1.2] Setting a record field to a variable:
\[
  \textcolor{gray}{xx :=}  \underbrace{accountRecord'balance}_{RecField} 
\]

\item[Case 2]  Attribution between structures:
\[
  \underbrace{newRecord}_{Record} :=  \underbrace{accountRecord}_{Record} 
\]
\end{description}

A B code example:

\begin{pascalcode}
xx := rec(id:1, balance:10);
xx`balance := xx`balance-1;
\end{pascalcode}

Lines 1 and 2 are represented in translation in LLVM code:

\begin{llvmcode}
%1 = getelementptr %struct.R_1* @Rec__account, i32 0, i32 0
store i32 1, i32 * %1
%2 = getelementptr %struct.R_1* @Rec__account, i32 0, i32 1
store i32 10, i32 * %2

%3 = getelementptr %struct.R_1* @Rec__account, i32 0, i32 1
%4 = load i32* %3
%5 = sub nsw i32 %4, 1
%6 = getelementptr %struct.R_1* @Rec__account, i32 0, i32 1
store i32 %5, i32* %6
\end{llvmcode}

The $RecField$ rule gets an element value from the record. The same rule is 
used on the left side of attribution as an $LValue$ and right side of 
attribution as an $Expression$. The functions \SBPos(n) and \SBType(n)  get 
respectively the position and type of $n$ element in the structure.

\begin{align*}
\begin{split}
  \trad{n}{\B{RecField}} \isdef \\
   &\LET \ell = \lbl  \AND  t = n.type \AND v = \B{head}(n.element) \AND\\
   &code1 \BE \llvm{"\PH{\ell} = getelementptr \PH{t}* \PH{v}, i32 0, i32  \SBPos(\B{n})\nl"} \IN \\
   &\IF \B{n.element} \mbox{ is \B{BOOL or NAT or INT} \AND n.element.length=1}\\
   &\quad \THEN  code1 \sep  \ell  \sep \SBType(\B{n})\\
   &\IF \B{head(n.element)} \mbox{ is \B{RecField} }  \THEN \\
   &\LET n2 \BE new \B{RecField}() \AND  \B{n2.base}  \BE \B{head}(n.element)\\
   &\AND \B{n2.element} \BE \B{tail}(n.element) \AND \\
   &\quad code2, label, pos = \trad{n2}{RecField} \IN\\
   &\quad \llvm{"\PH{code1} \PH{code2}"} \sep  label  \sep pos \\
\end{split}
\end{align*}

The attribution from record structure to a record structure variable, Line 1 
from the last B code example, is translated by a sequence of applications of
the  rule $RecField$. This rule was also designed to support nested structures.

The structure attribution are made in atomic way and it is simple example:

\begin{pascalcode}
simpleRec := rec(elem1 : 2, elem2:3);
simpleRec := rec(elem1 : simpleRec`elem1+1, elem2: simpleRec`elem1+1);
\end{pascalcode}

This code is translated to the following code and solved in three steps: solve
the expression, store it to a temporal variable and  store the result in the
final place.

\begin{pascalcode}
VAR tmp1, tmp2 IN
tmp1 := 1;
tmp2 := 2;
simpleRec`elem1 := tmp1;
simpleRec`elem2 := tmp2;
END;
VAR tmp3, tmp4 IN
tmp3 := simpleRec`elem1 + 1;
tmp4 := simpleRec`elem1 + 1;
simpleRec`elem1 := tmp3;
simpleRec`elem2 := tmp4;
END
\end{pascalcode}

In this example, note that the final value of $simpleRec'elem2$ is 2. This 
happens due to atomic assignment. This translation is represented by the 
$Record$ rule and it is omitted, because it is meta rule very simple.

The \textbf{case 2} is the atrribution between records:

\begin{pascalcode}
  newSimpleRec = simpleRec;
\end{pascalcode}

The following code is its translation to LLVM:

\begin{llvmcode}
  %4 = bitcast %struct.database* %newSimpleRec to i8*
  %5 = bitcast %struct.database* %simpleRec to i8*
  call void @llvm.memcpy.p0i8.p0i8.i64(i8* %4, i8* %5, i64 12, i32 0, i1 false)
....
; Function Attrs: nounwind
declare void @llvm.memcpy.p0i8.p0i8.i64(i8* nocapture, i8* nocapture readonly, i64, i32, i1) #1
\end{llvmcode}

The instruction bitcast is used to get the reference pointer from source address 
and the destination address. The function \textit{memcpy} get a sequence of argument.
The first argument is a pointer to the destination, the second is a pointer to the source.
The third argument is an integer argument specifying the number of bytes to copy, the
fourth argument is the alignment of the source and destination locations, and the fifth
is a boolean indicating a volatile access. By default, the last two arguments are 0 and false.

\subsubsection{Arrays}

The arrays store a fixed-size sequential collection of elements with the same type. 
They consist of contiguous memory locations, where the lowest address (0) 
corresponds to the first element and the highest address to the last element.
In B, a concrete array is a total function where the domain indexes the range 
elements.

There are three attribution cases related to array.
\begin{description}

\item[Case 1.1] - Initializing a unique value to an array using Cartesian product:
\[
\textcolor{gray}{arr :=} \underbrace{\{\underbrace{\underbrace{0..2 \times 0..2 \times 0..3  }_{seq\ Exp} \times \underbrace{\{5\}}_{Exp} }_{Product}\}}_{Array}
\]

\item[Case 1.2] - Initializing different values to an 
array\footnote{B0 definition requires the declaration of all values.} using a sequence of map:
\[
\textcolor{gray}{arr:=} \underbrace{\{\underbrace{(\underbrace{(0 \mapsto 0) \mapsto  0)}_{seq Exp} \mapsto \underbrace{1}_{Exp} , ... ,
(\underbrace{(2 \mapsto 2) \mapsto  3)}_{seq\ Exp} \mapsto \underbrace{27}_{Exp} }_{seq\ Map}\}}_{Array}
\]

\begin{comment}
\$VarName = global [size X type] X [(type value,)^+]

|array|Type = LET
				  IF array is Product THEN dec = |array.arg1, array.arg2 |CProduct  ( Quebrando o padrão?)
				  ELSEIF array is seq Map THEN dec = |array.Map |Map
				IN	  
					" global "^ dec
				END
|sIntl, value |CProduct = 
    	IF sIntl is  empty THEN  ""
	ELSE THEN
		LET sizeType = |sIntl, "i32"|PrintSizeType &
		 tailSizeType=|sIntl.tail,"i32"|PrintSizeType & 
		 values= |tailSizeType,values|CProduct IN
		"[sizeType]  [ tailSizeType  values ]"		
		END
	END

|sIntl,t| PrintSizeType =
    	IF sIntl.size == 1  THEN  "t"
	ELSE THEN
		LET size = sIntl.head.size()  &
			type = |sIntl.tail,t| PrintSizeType    &
		"[size X type ]"		
		END
	END

|array|Type = IF t is Map THEN ...
\end{comment}

\item[Case 2.1] - Updating a selected element from array by a variable value :
\[  \underbrace{\underbrace{arr}_{Array}(\underbrace{1+ii,1+jj,1+xx}_{seq\ Exp})}_{ArrayItem} \textcolor{gray}{:= xx}
\]
\item[Case 2.2] - Setting a selected element from array to a variable:
\[ 
\textcolor{gray}{xx :=} \underbrace{\underbrace{arr}_{Array}(\underbrace{1+ii,1+jj,1+xx}_{seq\ Exp})}_{ArrayItem}
\]
\item[Case 3] - Attribution between arrays\footnote{B0 definition requires the same type.}:
\[
\underbrace{arr1}_{Array} := \underbrace{arr2}_{Array} 
\]
\end{description}

\subsubsection{Relation arrays B0 and LLVM}

The arrays in B0 are represented by functions where an index points to an 
element. The different ways to evaluate an array are described as follows:

\subsubsection{Initialing an array}
The following two attribution commands have the same semantic effect.

\begin{pascalcode}
array2 := ( 0..1 * 0..1 * 0..3 * {5});

array2 := { 0|->0|->0|->5 , 0|->0|->1|->5 , 
            0|->1|->0|->5 , 0|->1|->1|->5 , 
            1|->0|->0|->5 , 1|->0|->1|->5 , 
            1|->1|->0|->5 , 1|->1|->1|->5 ,
            0|->0|->2|->5 , 0|->0|->3|->5 ,
            0|->1|->2|->5 , 0|->1|->3|->5 , 
            1|->0|->2|->5 , 1|->0|->3|->5 , 
            1|->1|->2|->5 , 1|->1|->3|->5 };
\end{pascalcode}

An alternative way to represent these attributions cannot be by
using sequential attributions as follow:

\begin{pascalcode}
array2(0,0,0) := 5;
array2(0,0,1) := 5;
// ... - A sequence of attributions
array2(1,1,3) := 5;
\end{pascalcode}

It is not possible because B array attributions are atomic. The B array 
attribution is made in two steps. First, the sequence of new value is calculated and
copied to its temporal place. Second, the sequence of new values is setted 
in its address. Finally, considering these steps and the execution of
following code then the value of ``array2(0,0,1)'' is ``2''.

\begin{pascalcode}
 array2 := ( (0..1) * (0..1) * (0..3) * {1});
 array2 := { 0|->0|->0|->2 , 0|->0|->1|->(array2(0,0,0)+1), ...  };
\end{pascalcode}

The LLVM instruction \textit{getelementptr} is used to select an element in
the array or structure. A simple example is:

\begin{llvmcode}
%7 = getelementptr [2 x [2 x [4 x i32]]]* @array2, i32 0, i64 1, i64 1, i64 3
\end{llvmcode}

The first two parameters define the type and name of the variable, and the
others are indexes to select the element in the data type. The third parameter (i32 0)
selects the first element in the pointer (* @array2). The last elements 
(i64 1, i64 1, i64 3) select the value from position (1,1,3) storing it in \%7.
This array is multidimensional, so several indexes must be declared separated
by comma to get the specific element.

\subsubsection{Setting and getting array}

The following blocks illustrate a B code setting a value to array
and its translation to LLVM.

\begin{pascalcode} 
VAR xx, yy, zz IN
	xx=1;
	yy=1;
	zz=3;
	array2(x,y,z):=7
END
\end{pascalcode}

This example is translated in four steps. First, the variables \%x, \%y, \%z 
are allocated. Second, the respective values 0, 1, 3 are stored in variables. 
Third, the variables are loaded to local variables and converted to 64 bits. 
Finally, the address of array is calculated in local variable \%7 and the value 
7 is stored in it.

\begin{llvmcode}
%xx = alloca i32
%yy = alloca i32
%zz = alloca i32
store i32 1, i32* %xx
store i32 1, i32* %yy
store i32 3, i32* %zz
%1 = load i32* %zz
%2 = sext i32 %1 to i64
%3 = load i32* %yy
%4 = sext i32 %3 to i64
%5 = load i32* %xx
%6 = sext i32 %5 to i64
%7 = getelementptr[2x[2x[4xi32]]]*@array2,i32 0,i64 %6,i64 %4,i64 %2;getting an element
store i32 7, i32* %7         ;setting an element
\end{llvmcode}

\subsubsection{Rules for support array}

\paragraph{Product} declares all address and a unique value from array, it is represented by \textbf{case 1.1}.
The first case uses a declared constant that is attributed to the variable:
\begin{llvmcode}
@array = global [3 x [3 x [4 x i32]]] 
[[3 x [4 x i32]] 
	[[4 x i32] [i32 5, i32 5, i32 5, i32 5], 
	 [4 x i32] [i32 5, i32 5, i32 5, i32 5], 
	 [4 x i32] [i32 5, i32 5, i32 5, i32 5]], 
 [3 x [4 x i32]] 
	[[4 x i32] [i32 5, i32 5, i32 5, i32 5],
	 [4 x i32] [i32 5, i32 5, i32 5, i32 5],
	 [4 x i32] [i32 5, i32 5, i32 5, i32 5]],
 [3 x [4 x i32]] 
	[[4 x i32] [i32 5, i32 5, i32 5, i32 5],
	 [4 x i32] [i32 5, i32 5, i32 5, i32 5], 
	 [4 x i32] [i32 5, i32 5, i32 5, i32 5]]], align 16
\end{llvmcode}

The rule for declare array constants:

\begin{align*}
\begin{split}
  \trad{array}{\B{Type}} \isdef \\
   & \quad \IF \B{array} \mbox{ is \B{Product} } \THEN \\
   & \quad \quad \llvm{" global \PH{\trad{\llvm{"i32"}, array.arg1, array.arg2}{\B{Product}}}"}\\
%   & \quad \ELSIF array \mbox{ is \B{seq Map} } \THEN\\
%   & \quad \quad  \trad{\llvm{"i32"}, array.arg1, array.arg2}{\B{Map}} \\
\end{split}
\end{align*}

\begin{align*}
\begin{split}
  \trad{type, seqSet, value}{\B{Product}} \isdef \\
   & \LET\\
   & \quad \B{times} \BE \B{head}(seqSet).end - \B{head}(seqSet).start + 1\\
   & \quad left \BE \trad{seqSet,type}{ArrayType} \IN \\
   & \IF \B{seqSet.length} == 1 \THEN  \\
   & \quad \LET right = \trad{times,type,value}{\B{TV}} \IN\\
   & \quad \llvm{"\PH{left} [\PH{right}]"}  \\
   & \ELSE \\
   & \quad \llvm{"\PH{left} \PH{\trad{type,tail(seqSet),value}{\B{Product}}}"}\\
\end{split}
\end{align*}

\begin{align*}
\begin{split}
  \trad{times, type, value}{\B{TV}} \isdef \\
   &\LET res1 \BE \llvm{" \PH{type} \PH{value} "} \IN\\
   & \IF times == 1 \THEN \llvm{"\PH{res1}"} \\
   & \ELSE\\
   & \quad \llvm{"\PH{res1} , \PH{\trad{times-1,type,value}{\B{TV}}}"} \\
\end{split}
\end{align*}

\paragraph{seq Map} declares every address and values from array, it is represented by \textbf{case 1.2}.

The attribution using an sequence of Map are made in atomic way and it is simple example.

\begin{pascalcode}
 array2 := ( (0..1) * (0..1) * (0..3) * {1}); // Cartesian product
 array2 := { 0|->0|->0|->2, 0|->0|->1|->(array2(0,0,0)+1), ...}; // sequence of map
\end{pascalcode}

This code is translated in the following code and solved in three steps: solve 
the expression, store it to a temporal variable and  store the result in the 
final place.

\begin{pascalcode}
VAR tmp1, tmp2 IN
tmp1 := 2;
tmp2 := array(0,0,0)+1;
array(0,0,0) := tmp1;
array(0,0,1) := tmp2;
END;
\end{pascalcode}

In this example, note that the final value of $array(0,0,1)$ is 2. This 
happens due to atomic assignment. This translation is represented by the 
$Map$ rule and it is omitted, because it is meta rule very simple.

\paragraph{Array item} selects an item in expressions as ``$a(n1,...,n_{n})$'', 
where $a$ is $n.base$ and $n1,...,n_{n}$ is a sequence of indexes that 
calculates the address of element. This rule also supports to translate 
expressions of arrays. This rule is used to support array attributions 
from \textbf{cases} \textbf{2.1} and \textbf{2.2}.

\begin{align*}
\lefteqn{\trad{n}{ArrayItem} \isdef} \\
&  \LET p_1 \sep v_1 \sep t_1  = \trad{n.base}{Exp}  \IN \\
&  \LET r \sep p_2  = \trad{n.indexes}{LRExp}  \IN \\
&  \LET v =  \variable \AND  t = \trad{\B{n.type}}{Type}  \IN \\
&  \llvm{"\PH{p_1} \PH{p_2} \PH{v} =  getelementptr \PH{t_1}, \PH{v_1}, i32 0 \PH{r}\nl"}  \sep v \sep t\\
\end{align*}

A list of type and variable is generated by the following rule.
\begin{align*}
&\trad{n}{LRExp} \isdef \LET p \sep v \sep t  = \trad{n}{ShiftExp}  \IN  \llvm{",\PH{t} \PH{v} "} \sep p\\
&\trad{n l}{LRExp} \isdef \LET r_1 \sep p_1 = \trad{n}{LRExp} \AND
 \LET r_2 \sep p_2 = \trad{l}{LRExp} \IN\\
& \quad  \llvm{"\PH{r_1} \PH{r_2}"} \sep \llvm{"\PH{p_1} \PH{p_2}"}\\
\end{align*}

The element access of array can need a shift in indexes. When the domain of 
array is an interval and it is not started from zero, it must adjust the 
indexes shifting the access, because the addressing  of array in intermediate 
representation LLVM starts at 0. To access an element from array the formula is 
$address(e(i1,..,in))  = i1-i1\_start, ... , in-in\_start$. The function $TSymbol$
returns the ``start'', ``end'' and type of the one index. The following rule 
makes the shift for each index of array.

\begin{align*}
&\trad{n}{ShiftExp} \isdef \LET p \sep v \sep t  = \trad{n}{Exp} \AND \\
&\quad iStart \sep iEnd \sep iType = \trad{n}{TSymbol}  \AND  \PH{\ell} = \lbl \IN \\
&\quad \IF ( iStart == 0 ) \THEN  \llvm{",\PH{t} \PH{v} "} \sep p \\
&\quad \ELSE \LET  p' \sep v \sep v' \sep t'  = \trad{t,v,itype,istart}{Cast} \IN \\
&\quad \llvm{",\PH{t} \PH{\ell} "} \sep  \llvm{"\PH{p} \PH{p'}  \PH{\ell} = sub \PH{t'} \PH{v}, \PH{v'}\nl"}  \\
\end{align*}

The \trad{}{ShiftExp} rule does a subtraction and this operation must have operators with the same data types so the following cast rule is used:

\begin{align*}
&\trad{t,v,t',v'}{Cast} \isdef \LET  \ell = \lbl \IN \\
&\quad \IF ( t == t') \THEN p \sep v \sep v' \sep t  \\
&\quad \ELSE \IF (t<t') \THEN \\
&\quad \quad \llvm{"\PH{\ell}= bitcast \PH{t} \PH{v} , \PH{t'}\nl"}  \sep l \sep v' \sep t'\\ 
&\quad \ELSE \\ 
&\quad \quad \llvm{"\PH{\ell}= bitcast \PH{t'} \PH{v'} , \PH{t}\nl"} \sep v \sep l \sep t \\ 
\end{align*}

The \textbf{case 3} is the atrribution between arrays. This translation is 
very similar to atribution between records.

\begin{pascalcode}
	newArray = simpleArray;
\end{pascalcode}

The following code is its translation to LLVM:

\begin{llvmcode}
%x4  = bitcast [10 x i32]* @newArray to i8*
%x5  = bitcast [10 x i32]* @simpleArray to i8*
call void @llvm.memcpy.p0i8.p0i8.i64(i8* %x4, i8* %x5, i64 40, i32 0, i1 false)
\end{llvmcode}

\paragraph{Relevant translation restrictions in B2LLVM:}
\begin{itemize}
\item  B0 definition - An array must be initialized declaring all values 
to avoid no determinism.
\item B0 definition from CommenC - The operator overwrite from arrays is not supported, for example:

\begin{pascalcode}
array := array <+ {1|->2|->0|->4,  1|->2|->1|->5}
\end{pascalcode}

\end{itemize}