### 2014-08-22 DOCUMENTATION/SLIDES/20140422-LORIA/figures_minted/with_trace.tex

\documentclass[10pt,a4paper]{article}
\include{codeFormat}

\begin{document}

\begin{llvmcode}
;;1 The type for the state of "counter" is defined in "counter_i",
;; it is an aggregate such that:
;;1.1 Position "0" represents variable "value".
;;1.2 Position "1" represents variable "error".
%counter$state$ = type {i32, i1}
;;2 The type for references to state encodings of "counter" is:
%counter$ref$ = type %counter$state$*
;;3 The function implementing initialisation for "counter" is
;;  named "@counter$init$" and has the following parameters:
;;3.1 "%self$": address of LLVM aggregate storing state of "counter";
define void @counter$init$(%counter$ref$ %self$) {
\end{llvmcode}

\end{document}